### 2021-11-24 Scripts/Initial_Analysis.Rmd
FIxed top 5 ssr map

---
title: Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for
  Policy Response Under Climate Change
author: "Juliano Palacios Abrantes"
date: "09/03/2020"
output:
  word_document: 
    keep_md: yes
  html_document: default
subtitle: Initital analysis
editor_options: 
  chunk_output_type: console
---

```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}

library(MyFunctions)

#### Project's Library
packages <- c(
  "dbemImport",
  "tidyverse",
  "here", # for dbem_import `here()`da
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  "geosphere", # estimate distances between points `distm()`
  "ggrepel",
  "zoo", # for average mean
  "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "zeallot", # for Juanito's map
  "gmt", # for estimating distances between points
  "viridis",
  "cowplot",
  "rfishbase",
  "stargazer", # For nice lm tables
  "tidytext", # for boxplot order
  "moments", # for skewness and kurto
  "pgirmess" # for non parametric post hoc test
)

my_lib(packages)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

```

# Methods

See methods section of the paper (e.g. Manusript.Rmd)

## Creating Data

Here we re-arrange existing data for the analysis. We have three steps:

1- Set the SAU relations between index and EEZs and include lat long data for each index

2.- Deteremine the centroid of each EEZ and asigned an id to each neighbouring interaction (e.g. Chile and Peru or Canada West Coast and US West Coast)

3.- Set Neigbouring IDs based on the interactions between neighbouring EEZs for all of those sharing transboundary species

```{r pre_data_creation, eval = F, echo = F}

## ---------------##
#### 1. SAU relations between INDEX and Country's EEZs
# Re-ran with new names data
## ---------------##

# EEZIDs_List <- my_path("G", extra_path = "Spatial/V_Lam", name="Updated_EEZList_17June2016.xlsx", read = TRUE) %>%
# clean_names()

EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
  clean_names()
colnames(EEZ_CellID) <- c("eezid","index")

# SAU name relatioal table (Fish For Visa)
sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="matching_names.csv", read = TRUE)

Lon_Lat_DBEM <- my_path("G", extra_path = "Spatial", name="Lon_Lat_DBEM.txt", read = TRUE, header = FALSE)
colnames(Lon_Lat_DBEM) <- c("index","lon","lat")

# Index_Code <- EEZIDs_List %>% 
#   left_join(EEZ_CellID) %>% 
#   rename(eez_name = name) %>% 
#   left_join(Lon_Lat_DBEM,
#             by = "index")

Index_Code <-  sau_clean_names %>% 
  select(eezid = sf_eezid) %>% 
  left_join(EEZ_CellID,
            by = "eezid") %>% 
  left_join(Lon_Lat_DBEM,
            by =  "index")

# Save Index data for future 
# write_csv(index_code,
# paste(my_path("G",extra_path = "Spatial/DBEM"),"sau_index_code.csv",sep="")
# )

## ---------------##
#### 2. entroids 
## ---------------##

# Load SAU shapefile name and paths

# The path
Path_SAU <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")

# The File
File_Name <- "SAUEEZ_July2015.shp"

# Load it!
SAU_eez_sf <- st_read(dsn = Path_SAU,
                      layer =file_path_sans_ext(File_Name)
) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)

# Get the center poligon of each EEZ for map source/target with st_centroid
coords <- as.data.frame(st_centroid(SAU_eez_sf)) %>% 
  ungroup() %>% 
  select(eez_name,geometry) %>% 
  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
  # Manually fix issues
  mutate(
    longitude = ifelse(
      eez_name == "Fiji", 178.065033,
      ifelse(eez_name == "Tuvalu",177.6493,
             ifelse( eez_name == "New Zealand",174.8860,
                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,
                             ifelse( eez_name == "Brazil",-36.937,longitude
                             )
                     )
             )
      )
    )
  )

# Remove "c()"
coords$longitude <- gsub("\\(","",coords$longitude)
coords$longitude <- gsub("c","",coords$longitude)
coords$latitude <- gsub(")","",coords$latitude)

coords <- coords %>% 
  mutate_at(vars(longitude,latitude),as.numeric)

# Double check coords
# Note that Australia and Ggreenland have points in the middle of the country because of the EEZ shape

coords_map <- ggplot() +
  geom_sf(data = SAU_eez_sf, aes(), fill = NA) +
  geom_point(data = eez_centroid,
             aes(
               x = longitude,
               y = latitude
             )
  )

ggsave("coords_map.png",
       plot = coords_map,
       width = 10,
       height = 8,
       units = "in"
)

# Save dataframe so we don't need to run this ewvery time
# write_csv(coords,
#           "eez_centroids.csv")

## ---------------##
#### 3. Set Neigbouring IDs
## ---------------##

# Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE)

# Get id of neighbouring pairs
Neighbours_spp <- Transboundary_spp %>% 
  group_by(eez_name,eez_neighbour) %>% 
  summarise(n=n()) %>% 
  ungroup() %>% 
  select(-n) %>% 
  rownames_to_column("neighbour_id")

# Split dataframes to merge neighbouring ids
df1 <- Neighbours_spp
df2 <- Neighbours_spp
colnames(df2) <- c("neighbour_id", "eez_neighbour", "eez_name")

# Merge neighbouring ids
df_full <- bind_rows(df1, df2) %>% 
  group_by(eez_neighbour,eez_name) %>% 
  summarise(neighbour_id = min(neighbour_id))

# Save dataframe so we don't need to run this every time
# write_csv(df_full, 
#           "Neighbours_eez_id.csv")

```

## Estimating Proportion Change

This part of the analysis estimates the Stock Share Ratio and its change under cliamte change under two time frames; early (2030) and mid (2050). Note that 2020 is the average of 2021 to 2040 and mid is the average of 2041 to 2060.

### Fun `EstPropChange`

```{r EstPropChange, eval = F, echo = F}

EstPropChange <- function(taxon_key,
                          year,
                          hist_y = 2005,
                          data_type = "Catch",
                          ensemble_list, # expects a list of ensembles to averrage by
                          path,
                          neighbours){

  # Loads species data for all ensembles
  Spp_Dist <- bind_rows(
    lapply(
      ensemble_list,
      read_dbem_ens,
      taxon_key = taxon_key,
      year = year,
      data_type = data_type,
      model = "GFDL",
      rcp = "85",
      # path = "/Volumes/DATA/DATA",
      my_path = F
    )
  )

  # Filter Species distribution to that of the neighbouring EEZs
  # ------------ #
  # Warning message:
  # Column `eez_name` joining factor and character vector, coercing into character vector 
  # ------------ #

  suppressWarnings(
    Trans_EEZ <- Transboundary_spp %>% 
      filter(taxon_key %in% Spp_Dist$taxon_key) %>% 
      left_join(index_code,
                by = c("eez_name")
      )
  )

  # Get neighbouring id
  Neighbours_List <- Neighbours %>% 
    semi_join(Trans_EEZ,
              by = c("eez_name","eez_neighbour")
    )

  # Selecting only the transboundary nature of the species
  Trans_Spp <- Spp_Dist %>%
    filter(index %in% Trans_EEZ$index) %>% 
    left_join(index_code,
              by = "index") %>% 
    filter(eez_name %in% Neighbours_List$eez_name)

  # -------------------- #
  # Estimate area index
  # The number of species' cells present within each country's EEZ
  # -------------------- #

  #Determines the amount of catch present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(taxon_key,
             eez_name,
             year,
             ensemble) %>% 
    summarise(n_spp_eez = length(unique(index)),
              total_mcp = sum(value)
    ) %>% 
    left_join(Neighbours_List,
              by = "eez_name")

  # Split dataframes to merge latter
  Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      taxon_key,
      Name=eez_name,
      n_spp_eez,
      total_mcp,
      year,
      ensemble
    )

  Neighbour_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      taxon_key,
      n_spp_eez,
      total_mcp,
      Name=eez_neighbour,
      eez_name,
      year,
      ensemble
    )

  # Estimate Area index and % change
  Per_Ghange_D <- full_join(Territory_T,
                            Neighbour_T, 
                            by = c("Name","taxon_key","year","ensemble")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T),
           MCP_Total = sum(total_mcp.x,total_mcp.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(eez_name = Name,
           eez_neighbour =eez_name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y,
           mcp_Country = total_mcp.x,
           mcp_Neighbour = total_mcp.y) %>% 
    mutate(area_index = n_spp_Country/Spp_Total,
           catch_proportion = mcp_Country/MCP_Total) %>% 
    # Determine the three time frames for comparrison
    mutate(
      time_step = ifelse(year < hist_y,"historical",
                         ifelse(year >= 2021 & year <= 2040, "early",
                                ifelse(year >= 2041 & year <= 2060, "mid","NA")
                         )
      )
    ) %>%
    filter(time_step != "NA") %>% 
    # Time mean and sd
    ungroup() %>% 
    group_by(taxon_key,ensemble,eez_name,eez_neighbour,time_step) %>% 
    summarise_at(vars("area_index","catch_proportion"),
                 c(temp_mean=mean,
                   tempo_sd=sd)
    ) %>% 
    # ensemble mean and sd
    group_by(taxon_key,eez_name,eez_neighbour,time_step) %>%
    summarise_at(vars("area_index_temp_mean","catch_proportion_temp_mean"),
                 c(ensemble_mean = mean,
                   ensemble_sd = sd
                 )
    )

  # --------------- #
  # End function
  # --------------- #

  if(nrow(Per_Ghange_D) > 0){

    # Save Result
    name <- paste("proportion_",taxon_key,".csv",sep="")
    path_name <- paste(my_path("R",extra_path = "Proportion_2005"),name,sep= "")

    write_csv(Per_Ghange_D,
              path_name)

    return(print(paste("analysis done",taxon_key)))
    # return(Per_Ghange_D)

  }else{
    print(paste("no result for",taxon_key))
  }

  # Or return a df
  # return(Per_Ghange_D)
}

EstPropChange(taxon_key = 600004,
              year = year_df$year,
              hist_y = 2000,
              data_type = "Catch",
              ensemble_list = c(102,103,104),
              # path = path,
              neighbours = neighbours)

```

### Proportion Control Panel

```{r Proportion_control_panel, eval = F, echo = F}

## ----------------- ##
# Data needed for estimating year of emergence
## ----------------- ##

# SAU relations between INDEX and Country's EEZs (see chunk 3)
index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)

# Neighbour list and their respective id (see chunk 3)
Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)

# Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE) %>% 
  clean_names() %>% 
  select(-v1)

## ----------------- ##
# Global variables
## ----------------- ##

# Year timeframe
year_df <- tibble(year =  c(seq(1951,2005,1), seq(2006,2099,1)),
                  time_period = c(rep("historic", 55), rep("future", 94))
)

# List of ensemble members
# ensemble_list <- seq(102,103,1) # for testing
ensemble_list <- seq(102,111,1)

#  List os species
spp_list <- unique(Transboundary_spp$taxon_key)

modeled_spp <- list.files(my_path("R", extra_path = "Proportion_n"))

modeled_spp <- gsub("proportion_","",modeled_spp)
modeled_spp <- gsub(".csv","",modeled_spp)

spp_list <- spp_list[!spp_list %in% modeled_spp]

```

### Proportion routine

```{r Proportion_routine, eval = F, echo = F}

####_______________________ ####
# Estimating proportion change
####_______________________ ####
gc()
t_start <- Sys.time()

mclapply(
  spp_list,
  EstPropChange,
  year = year_df$year,
  data_type = "Catch",
  ensemble_list = ensemble_list,
  path = DBEM_path,
  neighbours = Neighbours
)

t_end <- Sys.time()-t_start;t_end

# Hall 1000 ran it in 5.33 hours past
# Hall 1000 ran it in  hours (July 2020)

```

## Estimating emerging year

This rutine estimates the time (e.g. year) in which the changes in stock share ratio is more likely to happen based on historical variations (e.g. model noise) and future trends (e.g signal). This result uses teh average of the 10 ensemble members

### Fun `GetCentroids`

This function estimates the distance between the centroid of each neighbouring EEZ and the distribution of the species along both EEZs. We define species centroid as the tile with the highest value of maimum catch potential from all of the species distribution in each year.

```{r GetCentroids, eval = T, echo = F}

GetCentroids <- function(taxon_key,
                         year, 
                         data_type = "Catch",
                         ensemble, 
                         path,
                         Neighbours,
                         eez_centroid,
                         p # percentile of data for central distribution
){

  # ----------------- #
  # Get model data from spps for one ensemble
  #  ---------------- #
  if(ensemble == 102){
    print(taxon_key)
  }

  Spp_Dist <- read_dbem_ens(
    taxon_key = taxon_key,
    year = year,
    data_type = data_type,
    ensemble = ensemble, # change me to ensemble
    path = my_path("G","DBEM")
  )

  # ----------------- #
  # Extra step 
  # ----------------- #

  #### Creates a data to control computational time (lol!)
  #   if(file.exists(paste(my_path("R"),"spp_df.csv",sep="")) == FALSE){
  #   spp_df <- tibble() %>% 
  #     write_csv(.,
  #               paste(my_path("R"),"spp_df.csv",sep=""))
  # }
  # 
  # suppressMessages(
  #   spp_df <- my_path("R",name="spp_df.csv",read = TRUE)
  # )
  # spp_df <- tibble(taxon_key,
  #                  ensemble) %>% 
  #   bind_rows(spp_df) %>% 
  #   write_csv(.,
  #             paste(my_path("R"),"spp_df.csv",sep=""))

  # ----------------- #
  # Estimate distance between centroids
  # ----------------- #

  # Filter only the transboundary nature of the species 
  Trans_Spp <- Spp_Dist %>%
    left_join(index_code,
              by= "index") %>% 
    filter(!is.na(value)) %>% 
    semi_join(Transboundary_spp, # filter only transboundary cases
              by = c("taxon_key","eez_name")
    )

  # ----------------- #
  # Extra step 
  # ----------------- #
  # Controls for cases with no-data
  nr <- nrow(Trans_Spp)
  if(nr > 0){

    # Filter eez by those transboundary
    unique_eez <- unique(Trans_Spp$eez_name)

    Neighbours_combo <- Neighbours %>% 
      filter(eez_name %in% unique_eez,
             eez_neighbour %in% unique_eez)

    # Get the centroid of each country
    EEZ_centroid <- eez_centroid %>% 
      filter(eez_name %in% Trans_Spp$eez_name) %>% 
      mutate(taxon_key = taxon_key)

    # ----------------- #
    # Centroid function
    # ----------------- #

    # This function estimates the distance of each neighbouring EEZ
    # It uses Ids wich is a pre-determine identifier of neighbouring
    # EEZs. Fo example Peru and Chile have Id = 167, and viceversa

    CentroidsFx <- function(Ids){

      # Filter data by Ids
      EEZ_ids <- Neighbours_combo %>% 
        filter(neighbour_id == Ids)

      Sub_Trans_Spp <- Trans_Spp %>% 
        filter(eez_name %in% EEZ_ids$eez_name)

      Sub_EEZ_centroid <- EEZ_centroid %>% 
        filter(eez_name %in% EEZ_ids$eez_name)

      # Function to get percentiles
      # https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/
      quibble <- function(x, q = p) {
        tibble(x = quantile(x, q), q = q)
      }

      # Get the percentile treshold per year, ensemble and neighbour id
      percentile <- Sub_Trans_Spp %>% 
        left_join(EEZ_ids, by = "eez_name") %>% 
        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
        do(quibble(.$value, p))

      Centroids <- Sub_Trans_Spp %>% 
        left_join(EEZ_ids, by = "eez_name") %>% 
        left_join(percentile,
                  by = c("taxon_key","year","ensemble","neighbour_id")
        ) %>% 
        mutate(to_filter = ifelse(value < x,"out","keep")) %>% 
        filter(to_filter == "keep") %>% 
        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
        summarise_at(vars(lon,lat),mean) %>% 
        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
        # select(eez_name =eez_name.y, everything() ,-eez_name.x) %>% 
        # Estimate distances between spp centroid and EEZ centroid
        mutate(
          distance = geodist(longitude, # eez longitude
                             latitude,  # eez latitude
                             lon,  # species longitude
                             lat, # species latitude
                             units="km")
        ) %>% 
        select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)

      ### Code to produce graphs for Brazil and Uruguay

      # name <- paste(taxon_key, "distrbution p = 90%") 
      # Centroids %>% 
      #   # filter(year %in% c(1951,2000,2010)) %>% 
      #   ggplot() +
      #   geom_point(
      #     aes(
      #       x = lon,
      #       y = lat,
      #       size = distance,
      #       color = year,
      #       shape= eez_name
      #     )
      #   ) +
      #   geom_point(
      #     aes(
      #       x = longitude,
      #       y = latitude,
      #     ),
      #     fill = "red"
      #   ) +
      #   geom_sf(data = braur_sf,
      #           aes()
      #   )  +
      #   ggtitle(name)
      # 
      # ggsave(paste(taxon_key,"distribution.jpg",sep="_"),
      # plot = last_plot(),
      # width = 10,
      # height = 8,
      # units = "in",
      # path = my_path("D")
      # )

      return(Centroids)

    }

    # Determine the ids
    Ids <- unique(Neighbours_combo$neighbour_id)

    # Run sub-function
    Centroids_Data <- bind_rows(
      lapply(Ids,CentroidsFx)
    )

  }else{
    # In case there's no data
    Centroids_Data <- tibble()

  }

  return(Centroids_Data)

}

# Test function

# GetCentroids(taxon_key = 600004,
#              year = year_df$year[1:5],
#              data_type = "Catch",
#              ensemble = 102,
#              Neighbours = Neighbours,
#              eez_centroid = eez_centroid,
#              p = 0.95
#              )

# head(Test_centroids)
# unique(Test_centroids$ensemble)
# unique(Test_centroids$neigh)
#

# braur_sf <- World_ %>% 
#   filter(name %in% c("Brazil","Uruguay"))

# ggplot(braur_sf) +
#   geom_sf() +
# geom_point(data = subset(Centroids, year %in% seq(1951,2100,20)),
# geom_point(data = subset(Centroids,eez_name =="Brazil"),
# geom_point(data = Centroids,
#            aes(
#              x = lon,
#              y = lat,
#              color = year#,
#              # size = value
#            )
# ) #+
# facet_wrap(~to_filter+eez_name)

```

### Fun `TransIndex`

This functions takes on `GetCentroids` function and estimates the TransIndex (*TI*) for each pair of neighbouring countries. Presents a 10 years average of the results

$$TI =(\frac{D_{A,y}}{sd(D_{A,h})} - \frac{D_{B,y}}{sd(D_{B,h)}})^2$$

The year of emergece is defined as the first year where the trend is larger than the 95% CI of the historic variation.

Note: If no historic data available then emerging year is the first year of data in an EEZ

```{r TransIndex, eval = T, echo = T}

TransIndex <- function(Centroids, # Expecting the results from the GetCentroids function
                       Id, # A numeric value representing a neighbouring interaction (ONE!)
                       Hist_y = 2005,
                       Plot = F){ # The LAST year of Historical data

  # Filter data by neighbouring ID
  Centroids_id <- Centroids %>% 
    filter(neighbour_id %in% Id)

  EEZ_ids <- Neighbours %>% 
    filter(neighbour_id %in% Id)

  # Historical mean and sd of distance
  historical <- Centroids_id %>%
    filter(year <= Hist_y) %>% 
    ungroup() %>%
    group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
    summarise_at(vars("distance"),
                 c(hd_temp_mean=mean,
                   hd_temp_sd=sd,
                   n_yrs =~n())
    )

  # Future results
  future <- Centroids_id %>% 
    filter(year > Hist_y)

  # Get unique eezs
  Unique_EEZs <- unique(EEZ_ids$eez_name)

  # Transboundary Index
  Trans_index <- Centroids_id %>% 
    left_join(historical,
              by = c("taxon_key","eez_name","ensemble","neighbour_id")
    ) %>%
    # Re arrenging table for mutate
    gather("variable","value",distance:hd_temp_sd) %>%
    ungroup() %>% 
    mutate(value = ifelse(is.na(value),0,value)) %>% 
    mutate(
      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
      )
    ) %>%
    select(-eez_name,-variable) %>%
    spread(names,value) %>% 
    # mutate index
    mutate(
      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
      # )
    ) %>% 
    # Ten years average of results
    group_by(taxon_key,ensemble,neighbour_id) %>% 
    mutate(RMean = rollmean(x = trans_index, 
                            10, 
                            align = "right", 
                            fill = NA,
                            na.rm=T)
    ) %>% 
    # filter(year > 1960) %>% # drop first 10 years of running mean
    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)

  # Average of 10 years mean of ensemble members results
  mean_sd_index <- Trans_index %>% 
    group_by(taxon_key,year,neighbour_id) %>% 
    summarise_at(vars("RMean"),
                 c(trend=mean,
                   ensemble_sd=sd,
                   n_ens = ~n()

                 ),
                 na.rm=T
    )

  # Valid ensembles
  # Only perform analysis on ensembles containing all years between 1961 and 2040
  n_years <- mean_sd_index %>% 
    group_by(n_ens) %>% 
    summarise(
      n_year = n()
    ) %>% 
    filter(
      n_year == 149
    )

  mean_sd_index <- mean_sd_index %>% 
    filter(n_ens %in% n_years$n_ens)

  # Estimate noice from ensemble members
  historic_sd <- Trans_index %>% 
    filter(year <= Hist_y) %>% 
    # Alternative method sensitivity analysis
    # group_by(taxon_key,neighbour_id,year) %>% #s.d. historic data per year first
    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
    summarise_at(vars("RMean"),
                 c(trend = mean),
                 na.rm=T
    ) %>% 
    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
    summarise_at(vars("trend"),
                 c(hist_mean=mean,
                   noise=sd,
                   n_ens = ~n())
    )

  # For testing
  if(Plot == T & length(historic_sd$hist_mean) > 1){
    # Plot the result, bruh

    # Step to determine if data exists
    Plot_Test <- nrow(subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)))>0

    ggplot() +
      geom_line(data = mean_sd_index,
                aes(
                  x = year,
                  y = trend
                )
      ) +
      geom_ribbon(data = mean_sd_index,
                  aes(x=year,
                      ymin=historic_sd$hist_mean-historic_sd$noise,
                      ymax=historic_sd$hist_mean+historic_sd$noise),
                  alpha = 0.5,
                  fill = "grey50"
      ) +
      geom_ribbon(data = mean_sd_index,
                  aes(x=year,
                      ymin=historic_sd$hist_mean-(historic_sd$noise*2),
                      ymax=historic_sd$hist_mean+(historic_sd$noise)*2)
                  ,
                  alpha = 0.5,
                  fill = "grey30"
      ) +
      geom_line(data = mean_sd_index,
                aes(
                  x = year,
                  y = historic_sd$hist_mean
                ),
                colour = "grey50"
      ) +
      # Conditional, will only plot circles if the data exists
      {if(Plot_Test)geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
                               aes(
                                 x = year,
                                 y = trend
                               ),
                               colour = "red",
                               shape = 1,
                               size = 3
      )
      } +
      geom_line(data = Trans_index,
                aes(
                  x = year,
                  y = RMean,
                  color = as.factor(ensemble)
                ),
                alpha = 0.3
      ) +
      geom_vline(xintercept = Hist_y,colour="blue",alpha = 0.5) +
      scale_x_continuous(breaks = seq(1960,2100,5),limits = c(1960,2100)) +
      scale_color_brewer("Ensembles", palette = "Greens") +
      theme_classic()

    tk <- unique(mean_sd_index$taxon_key)

    ggsave(paste(tk,Id,"plot_example.jpg",sep="_"),
           plot = last_plot(),
           width = 10,
           height = 8,
           units = "in",
           path = my_path("R")
    )

  } #End test plot

  # If no historic data then emerging year is the first year of data
  if(nrow(historic_sd) == 0){

    Final_Result <- mean_sd_index %>%
      group_by(taxon_key,neighbour_id,n_ens) %>%
      summarise(emerging_yr= min(year)) %>%
      left_join(EEZ_ids,
                by = "neighbour_id") %>%
      ungroup() %>%
      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr) %>% 
      mutate(tresh= "NA")

    return(Final_Result)

  }else{

    Final_Result <- mean_sd_index %>%
      filter(year > Hist_y) %>%
      mutate(
        tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
        tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
      ) %>%
      gather("tresh","YesNo",tresh_one,tresh_two) %>% 
      filter(!is.na(YesNo)) %>% 
      group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
      summarise(emerging_yr= min(year)) %>%
      # summarise(emerging_yr= max(year)) %>% # Last year that treshold is crossed
      left_join(EEZ_ids,
                by = "neighbour_id") %>%
      ungroup() %>%
      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,tresh)

    return(Final_Result)
  }
}

```

### Fun `GetTransIndex`

This is the last function of the emerging year evaluation. It was made so we could run the analysis per each species. The function function simply runs the functions `GetCentroids` by ensabmble member and then `TransIndex` by pairing EEZ. Thus, with this function we run the whole set by species * paired EEZ * ten ensemble members.

```{r GetTransIndex, eval = T, echo = T}

GetTransIndex <- function(taxon_key,
                          year,
                          Hist_y,
                          data_type = "Catch",
                          path,
                          ensemble_list,
                          Neighbours,
                          eez_centroid,
                          p = NA,
                          Plot = F){
  t_start <- Sys.time()
  print("Starting Step One")
  # First step
  # Call Get centroids fucntion for centroids data per ensemble

  Step_one <- bind_rows(
    lapply(ensemble_list,
           GetCentroids,
           taxon_key = taxon_key, # for testing
           year = year_df$year,
           data_type = data_type,
           Neighbours = Neighbours,
           eez_centroid = eez_centroid,
           p = p
    )
  )

  # Set the different neighbouring pairs
  Ids <- unique(Step_one$neighbour_id)

  print("Starting Step Two")

  # Second Step

  ### TEST it 
  # for (i in 1:length(Ids)){
  #   print(Ids[i])
  #   print(TransIndex(Step_one,Ids[i]))
  # }

  # Get the transboudary index based on all centroids
  Step_two <- bind_rows(
    lapply(Ids,
           TransIndex,
           Hist_y = Hist_y,
           Centroid = Step_one)
  )

  name <- paste(my_path("R",extra_path = "/Emergence_95/"),taxon_key,"_emergence.txt",sep="")

  write_csv(Step_two,
            name)

  time_dif <- Sys.time()-t_start

  message <- paste("Analysis done for", taxon_key, "in",time_dif, "minutes")
  return(message)
}

# GetTransIndex(
#   taxon_key = 600076,
#   ensemble_list = ensemble_list,
#   year = year_df$year,
#   Hist_y = 2000,
#   data_type = "Catch",
#   Neighbours = Neighbours,
#   eez_centroid = eez_centroid
# )

# spp_list <- c(600004, 600245)
# 
#  test <- bind_rows(
#   lapply(spp_list,
#          GetTransIndex,
#          year = year,
#          data_type = "Catch",
#          path = path,
#          Neighbours = Neighbours,
#          eez_centroid = eez_centroid,
#          ensemble_list = ensemble_list
#   )
# )
#

```

### Emerging Control Panel

```{r Emerging_control_panel, eval = T, echo = T}

## ----------------- ##
# Data needed for estimating year of emergence
# Note: Use repo = F when running on the Beast or Hall1000
## ----------------- ##

# SAU relations between INDEX and Country's EEZs (see chunk 3)
index_code <- my_path("G", extra_path = "Spatial/SAU/", name="sau_index_code.csv", read = TRUE)

# EEZ list and their centroid of distribution (see chunk 3)
eez_centroid <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE)

# Neighbour list and their respective id (see chunk 3)
Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)

# Transboundary species and their sharing EEZs (see Palacios-Abrantes et al; FishForVisa)
Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE) %>% 
  clean_names()

# ------------------------------------------- #
### Estimate number of transboundary stocks ###
# ------------------------------------------- #

n_stocks <- Transboundary_spp %>% 
  left_join(Neighbours,
            by = c("eez_name", "eez_neighbour")) %>% 
  # filter(neighbour_id %in% c(1,105)) %>%
  # filter(neighbour_id == 470) %>%
  group_by(neighbour_id) %>% 
  summarise(n_stock = length(unique(taxon_key))) %>% 
  # View()
  group_by() %>% 
  summarise(t_trans_stocks = sum(n_stock)) %>% 
  pull(t_trans_stocks) # 9132

## ----------------- ##
# Global variables
## ----------------- ##

# Year timeframe
year_df <- tibble(year =  seq(1951,2099,1))

# Last year of "historic" data
Historic_year <- 2005

# Data path
# path = DBEM_path

# List of ensemble members
# ensemble_list <- seq(102,103,1) # for testing
ensemble_list <- seq(102,111,1)

#  List os species
spp_list <- unique(Transboundary_spp$taxon_key)

# Already modeled species
spp_done <- list.files(my_path("R", extra_path = "Emergence_alt_sens"))

spp_done <- gsub("\\_.*","",spp_done)

spp_reman <- spp_list[!spp_list %in% spp_done]

# For plotting test
rand_plot <- sample(spp_reman,5)

# For alternative method
alt_sens <- sample(spp_reman,30)

```

### Emerging Routine

```{r Emerging_routine, eval = F, echo = F}

####_______________________ ####
# Estimating year of eergence
####_______________________ ####
# n_cores <- 24
t_start <- Sys.time()

gc()

mclapply(spp_reman,
         GetTransIndex,
         year = year_df$year,
         Hist_y = Historic_year,
         data_type = "Abd",
         ensemble_list = ensemble_list,
         Neighbours = Neighbours,
         eez_centroid = eez_centroid,
         p = 0.95, # main is 95
         Plot = F
)

```

```{r Emerging_routine_DoPar, eval = F, echo = F}

library(doParallel)
library(doSNOW)

cl <- makeCluster(12)
registerDoSNOW(cl)

# To show processing times
repetitions <- length(spp_reman)
pb <- txtProgressBar(max = repetitions, style = 3)

progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

foreach(i=1:repetitions,
        .combine = rbind,
        .packages = c("MyFunctions",packages),
        .options.snow = opts,
        .verbose = T) %dopar% 
  GetTransIndex(
    taxon_key =spp_reman[i],
    ensemble_list = ensemble_list,
    year = year_df$year,
    Hist_y = 2000,
    data_type = "Catch",
    Neighbours = Neighbours,
    eez_centroid = eez_centroid
  )

close(pb)
stopCluster(cl)

```

## SST time of emergence

##### Fun `GetEnvData`

```{r get_env_data, eval = F, echo = T}

getEnvVar=function(yr,ensemble,variable){

  filepath <- paste("/Volumes/DATA/DATA/Environmental data/GFDL_CMIP5_Large_Ensemble/All var/ens",ensemble,"/",sep="")

  file = paste(variable,"_",yr,".txt",sep="")
  read_me <- paste(filepath,file,sep="")

  data <- bind_cols(lapply(read_me,FUN= data.table::fread, na.strings = "NA",header=F)
  ) %>% 
    mutate(index = seq(1,259200,1))

  colnames(data) <- c(as.character(ensemble),"index")

  data <- data %>% 
    gather("ensemble","variable",1:length(ensemble)) %>%  # fix it has to be untill 10
    mutate(variable = ifelse(variable == -9999.00000,NA,variable)) # Fixing NAs connotation

  # Retrun

  getTemp <- data %>% 
    mutate(ensemble = ensemble,
           year = yr)

  return(getTemp)
}

```

```{r temp_emergence, eval = F, echo = F}

TempEmer <- function(yrs, hist_y = 2005, variable, ensemble){ # The LAST year of Historical data

  # Load environmental data
  sst_data <- bind_rows(
    lapply(
      yrs,
      getEnvVar, 
      ensemble = ensemble, 
      variable = variable
    )
  ) %>% 
    left_join(index_code,
              by = "index") %>%  # join with EEZ indexes
    select(-lat,-lon,-eez_name) %>% 
    filter(!is.na(eezid)) %>% 
    # Get average temperature per eez
    group_by(ensemble,eezid,year) %>%
    summarise_at(vars("variable"),
                 c(hd_eez_mean=mean),
                 na.rm=T
    ) %>% 
    # estimate 10 years running mean
    group_by(ensemble,eezid) %>% 
    mutate(Rmean = rollmean(x = hd_eez_mean, 
                            10, 
                            align = "right", 
                            fill = hd_eez_mean)
    )

  # Average of 10 years mean of ensemble members results
  mean_sd_index <- sst_data %>% 
    group_by(eezid,year) %>% 
    summarise_at(vars("Rmean"),
                 c(trend=mean,
                   ensemble_sd=sd,
                   n_ens = ~n()

                 ),
                 na.rm=T
    )

  # Historic data
  historical <- sst_data %>%
    filter(year <= hist_y) %>%
    ungroup() %>%
    # get EEZ time average
    group_by(ensemble,eezid) %>%
    summarise_at(vars("Rmean"),
                 c(hd_temp_mean=mean),
                 na.rm= T
    ) %>% 
    # get ensemble member mean and noice
    group_by(eezid) %>% 
    summarise_at(vars("hd_temp_mean"),
                 c(hist_mean=mean,
                   noise = sd),
                 na.rm = T
    )

  # ----------------------- #
  # Plot for data viz
  # ----------------------- #

  # samp <- sample(mean_sd_index$eezid,30,replace = F)

  # mean_sd_index <- mean_sd_index %>% 
  # filter(eezid %in% samp)

  # plot_data <- mean_sd_index %>% 
  # left_join(historical,
  #           by = "eezid")

  # ggplot() +
  #   geom_line(data = plot_data,
  #             aes(
  #               x = year,
  #               y = trend
  #             )
  #   ) +
  #   geom_ribbon(data = plot_data,
  #               aes(x=year,
  #                   ymin=hist_mean-noise,
  #                   ymax=hist_mean+noise
  #                   ),
  #               alpha = 0.5
  #   ) +
  #   geom_line(data = plot_data,
  #             aes(
  #               x = year,
  #               y = hist_mean
  #               ),
  #             linetype = "dashed"
  #   ) +
  #   facet_wrap(~eezid,
  #              scales = "free") +
  #   # Conditional, will only plot circles if the data exists
  #   geom_point(data = subset(plot_data, trend > (hist_mean+noise) | trend < (hist_mean-noise)),
  #              aes(
  #                x = year,
  #                y = trend,
  #                colour = "red",
  #              ),
  #              shape = 1,
  #              size = 3
  #   ) +
  #   geom_vline(xintercept = hist_y,colour="black",alpha = 0.5) +
  #   scale_x_continuous(breaks = seq(1950,2100,5)) +
  #   theme_classic()

  # Future results
  future <- sst_data %>% 
    filter(year > hist_y) %>% 
    # get ensemble member mean and noice
    group_by(eezid,year) %>% 
    summarise_at(vars("Rmean"),
                 c(trend=mean,
                   f_noise = sd)
    ) %>% 
    group_by(eezid)

  Final_Result <- future %>%
    left_join(historical,
              by = "eezid") %>% 
    mutate(
      tresh_one = ifelse(trend > (hist_mean+noise) | trend < (hist_mean-noise),"One",NA),
      tresh_two = ifelse(trend > (hist_mean+(2*noise)) | trend < (hist_mean-(2*noise)),"Two",NA)
    ) %>%
    gather("tresh","YesNo",tresh_one,tresh_two) %>% 
    filter(!is.na(YesNo)) %>% 
    group_by(eezid,tresh) %>% 
    summarise(emerging_yr = min(year))

  return(Final_Result)
}

gc()

sst <- TempEmer(
  yrs = seq(1951,2070,1),
  hist_y = 2005,
  variable = "SST",
  ensemble = seq(102,111,1)
)

write_csv(sst,
          "ToE_sst.csv")

botTemp <- TempEmer(
  yrs = seq(1951,2070,1),
  hist_y = 2005,
  variable = "botTemp",
  ensemble = seq(102,111,1)
)

write_csv(botTemp,
          "ToE_bottemp.csv")

test <- sst %>% 
  filter(tresh == "tresh_one")

hist(test$emerging_yr)

gc()

test_map <- World_eez_sf %>% 
  left_join(test,
            by = "eezid") %>% 
  ggplot() +
  geom_sf(
    aes(
      fill = emerging_yr
    )
  )+ scale_fill_viridis(
    limits = c(2000,2070),
    breaks = seq(2000,2070,10)#,
    # direction = -1
  ) +
  my_ggtheme_m(leg_pos = "bottom") +
  ggsave("botTemp_ToE.png",
         width = 12,
         height = 10,
         units = "in"
  )

```

```{r william_code, eval = F, echo = F}

datascale<-subset(testdata,!is.na(MHW) & !is.na(MeanChange))
MHWBreaks<-c(0.2,0.4,0.5,0.6,0.8,2.1)
MeanChangeBreaks<-c(-0.4,0.6,0.7,0.8,1.0,2.0)
#testdata$MHW <- with(testdata, cut(MHW,
#                                   breaks=quantile(MHW, probs=seq(0,1, by=0.2), na.rm=TRUE),
#                                   include.lowest=TRUE))
testdata$MHW <- with(testdata, cut(MHW,
                                   breaks=MHWBreaks, na.rm=TRUE),
                     include.lowest=TRUE)
testdata$MHW <- as.numeric(testdata$MHW)
#testdata$MeanChange <- with(testdata, cut(MeanChange,
#                                          breaks=quantile(MeanChange, probs=seq(0,1, by=0.2), na.rm=TRUE),
#                                          include.lowest=TRUE))
testdata$MeanChange <- with(testdata, cut(MeanChange,
                                          breaks=MeanChangeBreaks, na.rm=TRUE),
                            include.lowest=TRUE)

testdata$MeanChange <- as.numeric(testdata$MeanChange)

library(ggplot2)
d <- data.frame(testdata,fill_val = rgb(blue=0.5,red = testdata$MHW/5, green = testdata$MeanChange/5))

world <- map_data("world")
map<- ggplot(d, aes(x=lon, y=lat)) +
  geom_tile(aes(fill = fill_val)) +
  scale_fill_identity()+
  geom_cartogram(data=world, map=world,
                 aes(x=long, y=lat, map_id=region),fill="white",color="black")+
  coord_quickmap(xlim=c(-180,180), ylim=c(-89, 90))+
  ylab("")+
  xlab("")+
  scale_y_continuous(expand = c(0, -5))+
  scale_x_continuous(expand = c(0, 0))+
  labs(title=paste("(A) Sea Surface Temperature",sep=""))+
  theme(axis.line = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position="bottom",
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank(),
        #axis.text.y=element_blank(),
        axis.text=element_text(size=9,colour="black"),
        plot.title=element_text(hjust=0,vjust=-0.4,face="bold",size=24)
  )
dlegend <- expand.grid(x = seq(0, 1, 0.25), y = seq(0, 1, 0.25)) %>%
  mutate(mix2 = rgb(red = x, green = y, blue = 0.5))
library(cowplot)
xyLegend <- ggplot(dlegend, aes(x, y)) +
  geom_tile(aes(fill = mix2)) +
  scale_fill_identity() +
  labs(x = "Higher MHW change->",
       y = "Higher mean change->") +
  theme_void() +
  theme(axis.title = element_text(size = 8),
        axis.title.y = element_text(angle = 90)) +
  coord_fixed()
p=ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(xyLegend, 0.05, 0.1, 0.2, 0.2)
#ggsave(p,filename="/Users/william/Google Drive/Projects/MHWs Global/MHW_SST_plot.png",height=9,width=8,dpi=300)
ggsave(map,filename="/Users/william/Google Drive/Projects/MHWs Global/MHW_SST_plot_map.png",height=8,width=11,dpi=300) (edited)

```

# Results

## Percentage Change

This results only shows the average percentage change in proportion of all species within an EEZ without relationship with neighbours

### Functions and Data needed

#### Fun `SummaryProp`

This function estimates the total change in MCP proportion of each species for each EEZ. It looses the connection with neighbours as you only know for x species how much is the proportion changing.

```{r fun_SummaryProp, eval = F, echo =F}

# taxon_key = "600147"
# top_stocks_per_country

SummaryProp <- function(taxon_key,links="NA", top = F){

  if(top == T){
    # Reads taxon data
    proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005c/"),"proportion_",taxon_key,".csv", sep="")) %>% 
      filter(!is.na(eez_neighbour)) %>% 
      rename(sf_eez_name = eez_name) %>% 
      left_join(Matching_names) %>% 
      semi_join(top_stocks_per_country) %>% 
      select(
      1:4,
      ensemble_mean = catch_proportion_temp_mean_ensemble_mean,
      ensemble_sd = catch_proportion_temp_mean_ensemble_sd) %>% 
      rename(eez_name = sf_eez_name)

  }else{

  # Reads taxon data
  proportion_data <- fread(paste(my_path("R",extra_path = "Proportion_2005c/"),"proportion_",taxon_key,".csv", sep="")) %>% 
    filter(!is.na(eez_neighbour)) %>% 
    select(
      1:4,
      ensemble_mean = catch_proportion_temp_mean_ensemble_mean,
      ensemble_sd = catch_proportion_temp_mean_ensemble_sd)

  }

  # Identify SSR treshold for natural spatial variability of the stock
  proportion_tresh <- proportion_data %>% 
    filter(time_step == "historical") %>% 
    mutate(
      top_tresh = ensemble_mean+(2*ensemble_sd), # two s.d.
      low_tresh = ensemble_mean-(2*ensemble_sd) # two.s.d
    ) %>% 
    select(taxon_key:eez_neighbour,top_tresh:low_tresh) %>% 
    left_join(proportion_data) %>% 
    # Only keeps change if surpasses treshold
    mutate(
      over_tresh = ifelse(time_step != "historical" & ensemble_mean > top_tresh | ensemble_mean < low_tresh,"keep",
                          ifelse(time_step == "historical","keep","drop")
      )
    ) %>% 
    # filter(time_step !="historical" | over_tresh == "keep") %>%
    filter(over_tresh == "keep") %>%
    select(-top_tresh,-low_tresh)

  if(links == "y"){
    # summarizes by country and neighbour
    country_summary <- proportion_tresh %>% 
      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
      summarise(
        mean_spp = mean(ensemble_mean, na.rm = T) # average spp change from all countries sharing
      ) %>% 
      filter(!is.na(mean_spp))

  }else{
    # summarizes by country
    country_summary <- proportion_tresh %>% 
      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
      summarise(
        mean_spp = mean(ensemble_mean,na.rm=T) # average spp change from all countries sharing that speceies
      ) %>% 
      filter(!is.na(mean_spp))
  }

  return(country_summary)
}

```

#### Fun `EEZProp`

The next step is to summarize all species per EEZ so our results shows average change of all species within an EEZ towards a neighbour.

**Note: Contains step to filter out changes less than +-2 s.d.**

```{r fun_EEZProp, eval = F, echo = F}

EEZProp <- function(taxon_list,links="NA",top = F){

  spp <- bind_rows(
    lapply(taxon_list,SummaryProp,links,top = top)
  )

  if(links == "y"){

    # Percentage of SSR stocks changing
    n_ssr <- spp %>% 
      group_by(eez_name,time_step) %>% 
      summarise(
        # n_spp = length(unique(taxon_key)),
        n_stock = n(),
      ) %>% 
      spread(time_step,n_stock) %>% 
      mutate_at(vars(early,mid),
                .funs = ~(./historical)*100
      ) %>% 
      select(-historical) %>%
      gather("time_step","n_per",early:mid)

    # Min, max and median change per eez_name and neighbor
    eez_summary <- spp %>% 
      group_by(eez_name,eez_neighbour,time_step) %>% 
      summarise(
        n_spp = length(unique(taxon_key)),
        n_stock = n(),
        min_eez = min(mean_spp, na.rm=T),
        max_eez = max(mean_spp, na.rm=T),
        sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
        mean_eez = median(mean_spp,na.rm=T)
      ) %>% 
      select(eez_name:max_eez,mean_eez,sd_eez,n_spp,n_stock) %>% 
      left_join(n_ssr)

  }else{

    eez_summary <- spp %>% 
      group_by(eez_name,time_step) %>% 
      summarise(
        n_spp = length(unique(taxon_key)),
        n_stock = n(),
        min_eez = min(mean_spp, na.rm=T),
        max_eez = max(mean_spp, na.rm=T),
        sd_eez = sd(mean_spp, na.rm=T), # for some weird reason it wont work if the order is fliped
        mean_eez = mean(mean_spp,na.rm=T)
      ) %>% 
      select(eez_name:max_eez,mean_eez,sd_eez,n_spp,n_stock)

  }

  return(eez_summary)

}

# SummaryProp(taxon_key)

```

#### Fun `PropChange`

Finally, we estimate the average proportion change of all species within each eez

```{r fun_PropChange, eval = F, echo =F}

PropChange <- function(taxon_list,links="NA", top = F){

  eez <- EEZProp(taxon_list,links, top = top)

  proportion_change <- eez %>% 
    gather("variable","value",n_spp:n_per) %>% 
    spread(time_step,value) %>% 
    mutate_at(vars(
      "per_early" = early,
      "per_mid" = mid
    ),
    funs(ifelse(variable == "n_per",.,round(((.-historical)/(abs(historical)))*100,2) # original
    )
    )
    )

  return(proportion_change)

}

```

### Run Results Analysis

#### Read Data for Analysis

```{r General_Data, eval = F, echo =F}

# --------------- #
# General Data needed
# --------------- #

# Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
Transboundary_spp <- my_path("D", extra_path = "Species/", name="Transboundary_spp.csv", read = TRUE)

# Get taxon list
Taxon_list_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))

# clean names for function
Taxon_list_emer <- gsub("_.*","",Taxon_list_emer)

# SAU exploited species list
Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
  clean_names() %>% 
  pull(taxon_name)

# Get habitat preference from fishbase
Spp_env_pref <- rbind(species(Exploited_spp_list, #species from rfishbase package
                              fields = species_fields$habitat[4])
) %>% 
  mutate(taxon_name = Exploited_spp_list) %>% 
  left_join(my_path("G","SAU","exploited_species_list.csv", read = T) %>%
              clean_names(),
            by = "taxon_name")

# Neighbor list and their respective id from Palacios-Abrantes et al 2020 (FishForVisa)
Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE)

Neighbours_n <- Neighbours %>% 
  group_by(eez_name) %>% 
  summarise(
    n = length(unique(eez_neighbour))
  )

# Clean names to match data
Matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T) %>% 
  rename(eezid = eez_id)

Fishing_names <- Matching_names %>% 
  group_by(sau_fishing_entity,fishing_entity_id) %>% 
  summarise(n=n()) %>% 
  select(-n)

# eez_centroids
Eez_centroids <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(Matching_names) %>% 
  select(eezid,latitude,longitude)

# --------------- #
# Load Proportion results
# --------------- #
# Get taxon list
Taxon_list <- list.files(my_path("R",extra_path = "Proportion_2005c"))

# clean names for function
Taxon_list <- gsub(".*_","",Taxon_list)

# clean names for function
Taxon_list <- gsub(".csv","",Taxon_list)

# Read proportion results
suppressMessages(
  Prop_results <- PropChange(Taxon_list, links="y") %>%
    filter(!is.na(eez_neighbour)) %>%
    mutate(per_mid_plot =ifelse(per_mid >= 100, 100,
                                ifelse(per_mid <= -100,-100,per_mid)
    ),
    per_early_plot =ifelse(per_early >= 100, 100,
                           ifelse(per_early <= -100,-100,per_early)
    )
    )
)
# For dat availability
# write_csv(Prop_results,
# "Data/ssr_data.csv")

# --------------- #
# Load year of emergence results
# --------------- #

Paths <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_emer,"_emergence.txt",sep="")

# Load all sp in one table
Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
  filter(
    n_ens == 10
  ) %>% 
  left_join(Neighbours,
            by = c("eez_name","eez_neighbour")
  ) %>% 
  rename(sf_eez_name=eez_name) %>% 
  left_join(Matching_names,
            by=  "sf_eez_name") %>% 
  clean_names() %>% 
  mutate(tresh = str_to_lower(tresh))

# warnings() for empty data species # New version of package being a pain in the arse

# Write data file for Github

# toe_data <- bind_rows(lapply(Paths, fread)) %>% 
#   filter(
#     n_ens == 10
#   ) %>% 
#   left_join(Exploited_spp_list) %>% 
#             select(taxon_key,taxon_name,common_name,everything(),-genus,-species)
# 
# write_csv(toe_data, "./Data/toe_data.csv")

# View(Eme_Res)

# --------------- #
# Load SAU catch and revenue data
# --------------- #

Sau_data <-fread(paste(my_path("D"),"SAU/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>%
  clean_names() %>% 
  # Remove high seas and 207 and 208 that are unknown
  filter(!fishing_entity_id %in% c(213,223)) %>%  # 207,208
  # Allocate catch of islands to UK
  mutate(fishing_entity_id = ifelse(fishing_entity_id %in% c(219,220), 183,fishing_entity_id)) %>% 
  # Convert to 2019 real USD
  mutate(values = 1.16 * value) %>%
  # Average 10 years of data
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise_at(c("catch","value"),
               mean,na.rm=T) %>% 
  gather("variable","value",catch:value)

# --------------- #
# Load shapefiles for figures
# --------------- #

# Load SAU shapefile name and paths

World_eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", "SAUEEZ_July2015.shp")) %>% 
  rename(eez_name = Name) %>% 
  # st_transform(crs = 4326) %>% # 4326
  st_transform(crs = "+proj=eck4") %>% # for tbular plot
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
  clean_names()

# World_eez_sf %>% 
#   filter(str_detect(eez_name,"Russia")) %>% 
#   ggplot() +
#   geom_sf() +
#   aes(fill = eez_name)

# Call world map from natural earth

World_map_dat <- Matching_names %>% 
  group_by(sau_fishing_entity,fishing_entity_id,iso_a3) %>% 
  summarise(n=n()) %>% 
  select(-n)

World_map <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf")) %>%
  clean_names() %>% 
  select(iso_a3,sovereignt,sov_a3,name,name_long,region_wb,continent,geometry) %>% 
  st_transform(crs = "+proj=eck4") %>%
  # st_transform(crs = 4326) %>% # 4326
  mutate(iso_a3 = ifelse(sovereignt == "Norway","NOR",
                         ifelse(name == "France","FRA",
                                iso_a3)
  )
  ) %>% 
  left_join(World_map_dat,
            by = "iso_a3")

# Load eez_centroid data
Eez_centroid <- my_path("D", "Spatial","eez_centroids.csv",read = T)

# Tranform to tubular
Eez_centroid_sf <-st_as_sf(Eez_centroid,
                           coords = c("longitude", "latitude")
) %>% 
  st_set_crs(4326) %>% # to match shapefiles
  st_transform(crs = "+proj=eck4")

# Transform sf to df
Eez_centroid_eck <- as.data.frame(Eez_centroid_sf)

Eez_centroid_eck <- do.call(rbind, st_geometry(Eez_centroid_sf)) %>%
  as_tibble() %>%
  setNames(c("lon","lat")) %>%
  bind_cols(Eez_centroid) %>%
  select(eez_name,longitude = lon,latitude = lat)

#

# See chunk Emerging_control_panel
n_stock <- 9132
n_eez <- length(unique(Eez_centroid_sf$eez_name))

### ------------ ##
# Select a treshold
selected_tresh <- "tresh_two"
### ------------ ##

```

### Centroid Supplemental Figure
```{r}

centroid_map <- ggplot() +
  geom_sf(data = World_eez_sf, aes()) +
  geom_sf(data = subset(World_map, !sovereignt %in% c("Antarctica")), aes(),fill = "grey90") +
  geom_sf(data = Eez_centroid_sf, aes()) +
  my_ggtheme_m(leg_pos = "")

ggsave(
  plot = centroid_map,
  "./Figures/centroid_map.png",
  width = 8,
  height = 5,
  units = "in"
)

```

###  Time of Emergence Analysis

```{r ToE_generals,  eval = F, echo = T}

# Total number of emerging stocks

# Global estimates
Eme_Res %>% 
  group_by(tresh,neighbour_id) %>% 
  summarise(
    emerging_stocks = length(unique(taxon_key))
  ) %>% 
  group_by(tresh) %>% 
  summarise(
    emerging_stocks = sum(emerging_stocks),
    prop_total = (emerging_stocks/n_stock)*100,
  )

# Global average ToE
Eme_Res %>% 
  group_by(tresh) %>% 
  summarise(
    # emerging_stocks = length(unique(taxon_key)),
    # prop_total = (emerging_stocks/9132)*100,
    mean= mean(emerging_yr),
    sd = sd(emerging_yr),
    min = min(emerging_yr),
    max = max(emerging_yr)
  )

# By 2030
Eme_Res %>% 
  filter(emerging_yr >= 2030) %>% 
  group_by(tresh,neighbour_id) %>% 
  summarise(
    emerging_stocks = length(unique(taxon_key))
  ) %>% 
  group_by(tresh) %>% 
  summarise(
    emerging_stocks = sum(emerging_stocks),
    prop_total = (emerging_stocks/n_stock)*100
  )

```

##### Fig. 2. Accumulative plots

Timeline of shifts in transboundary shared stock distribution. (A) Number of stocks (cumulative) whose distribution centroids that have shifted beyond natural variations over time (from 2006 to >2100), (B) Number of EEZs (cumulative) with the distribution centroid of at least one stock that have shifted beyond natural variations and (C) proportion of total shifting stocks and EEZs with at least one stock shifting beyond natural variability. Dashed lines represent the year by which countries have committed to reach full implementation of the 2030 Agenda, as well as mid and late change in range shift (Methods), respectively.

###### Generate data needed

```{r Time_lapse_data,  eval = F, echo = T}

# Per Stock
stock_d <-  
  Eme_Res %>% 
  left_join(Neighbours) %>% 
  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
  group_by(taxon_key,neighbour_id,tresh) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(tresh) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "stock"
  )

# max(stock_d$com_sum)

#  EEZ data
eez_d <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
  filter(n_ens == 10) %>% 
  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
  group_by(sf_eez_name,tresh) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(tresh) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "eez"
  )

# max(eez_d$com_sum)

# Plot Data
plot_data <- stock_d %>% 
  bind_rows(eez_d) %>% 
  filter(tresh != "NA") %>% 
  mutate(
    # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
    total = ifelse(level == "stock",n_stock,n_eez),
    per = com_sum/total*100
  ) %>% 
  gather("metric","value",com_sum,per) #%>% 
# filter(tresh == selected_tresh)

# Manually include rest of data until 2100
cont_eez <- plot_data %>% 
  filter(level == "eez") %>% 
  arrange(taxon_emer) %>% 
  slice(1:24) %>%
  select(-1) %>% 
  arrange(metric) %>% 
  mutate(
    value = ifelse(metric == "per",81,228),
    taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
  )

plot_data <- plot_data %>% 
  bind_rows(cont_eez) %>% 
  mutate(
    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs")
  )

dashed_lines <- tibble(
  x = c(2030
        # ,2050,2070),
  ),
  label = c("UN 2030 Agenda"
            # ,"Mid range shift   ","Late range shift   "
  )
)

# How many between 2006 and 2030?
plot_data %>% 
  filter(taxon_emer <= 2030,
         tresh == selected_tresh) %>%  # Presenting results for the second treshold
  group_by(level,metric) %>% 
  summarise(
    max(taxon_emer),
    min(taxon_emer),
    mean(taxon_emer),
    sd(taxon_emer)
  )

```

###### Plot histograms

```{r Time_lapse_plot, eval = F, echo = F}

### Three plots

count_eez_plot <- ggplot() +
  geom_area(data = subset(plot_data, metric == "com_sum" & level == "eez"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  )  +
  # Add dashed lines and labels
  geom_vline(
    data = dashed_lines,
    aes(
      xintercept = x
    ),
    linetype="dashed",
    color = "grey"
  ) +
  geom_text(data = dashed_lines,
            aes(
              x = x-4,
              y = 40,
              label = label
            ),
            size = 4,
            color = "white",
            angle = 90
  ) +
  scale_y_continuous("Cumulative number of EEZs",
                     expand = c(0,0)
  ) +
  scale_x_continuous("Year of range shift",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100")
  ) +
  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
  theme(
    axis.ticks.x = element_line()
  ) +
  scale_fill_viridis(discrete = T,
                     begin = 0.1, 
                     end = 0.1
  )

#### Count stock plot 
count_stock_plot <- ggplot() +
  geom_area(data = subset(plot_data, metric == "com_sum" & level == "stock"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  )  +
  # Add dashed lines and labels
  geom_vline(
    data = dashed_lines,
    aes(
      xintercept = x
    ),
    linetype="dashed",
    color = "grey"
  ) +
  geom_text(data = dashed_lines,
            aes(
              x = x-4,
              y = 750,
              label = label
            ),
            size = 4,
            color = "white",
            angle = 90
  ) +
  scale_y_continuous("Cumulative number of stocks",
                     expand = c(0,0)
  ) +
  scale_x_continuous("",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100")
  ) +
  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
  scale_fill_viridis(discrete = T,
                     begin = 0.7, 
                     end = 0.7
  ) + theme(
    axis.ticks.x = element_line()
  )

# Percentage plot
percentage_plot <- ggplot() +
  geom_area(data = subset(plot_data, level == "eez" & my_facet == "Percentage (%)"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  ) + 
  geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  ) + 
  # Add dashed lines and labels
  geom_vline(
    data = dashed_lines,
    aes(
      xintercept = x
    ),
    linetype="dashed",
    color = "grey"
  ) +
  geom_text(data = dashed_lines,
            aes(
              x = x-4,
              y = 14,
              label = label
            ),
            size = 4,
            color = "white",
            angle = 90
  ) +
  scale_y_continuous("Cumulative proportion of the total (%)",
                     expand = c(0,0)
  ) +
  scale_x_continuous("",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100")
  ) +
  my_ggtheme_p(leg_pos = "right", hjust = 0.5) +
  scale_fill_viridis(discrete = T,
                     begin = 0.1,
                     end = 0.7,
                     labels = c("EEZs","Stocks")
  )+theme(
    axis.ticks.x = element_line()
  )

# Legend
legend <- get_legend(percentage_plot)

percentage_plot <- percentage_plot + 
  theme(legend.position = "")

# Draw figure
plot_all <- ggdraw() +
  # Count stock plot
  draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
  # Count EEZ plot
  draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
  # Percentage plot
  draw_plot(percentage_plot, x = 0.61, y = 0 , width = 0.29, height = 0.90) +
  # legend
  draw_plot(legend, x = 0.84, y = 0 , width = 0.20, height = 1) +
  # Labels
  draw_plot_label(label = c("A", "B","C"), 
                  size = 20,
                  x = c(0.01, 0.285,0.59), 
                  y = c(1,1,1)
  )

# Save plot
save_plot(
  "./Figures/cumu_per_freq.png",
  plot_all,
  base_height = 5,
  base_width = 12
)

```

##### Fig. 3 Mapping the time of emergence and the participation

Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons display average year of range shifts within them. Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cold colors represent a late shift/low fishing revenue. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue. B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region. Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand.

###### Generate mean ToE across EEZs' stocks

```{r emergence_map_data, eval = F}
# Per EEZs
Eez_d_hist <- Eme_Res %>% 
  group_by(eezid,sau_fishing_entity,tresh) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(
    mean_yr = mean(emerging_yr,na.rm=T),
    sd_yr = sd(emerging_yr,na.rm=T),
    n_taxa = length(unique(taxon_key))
  ) %>% 
  gather("metric","value",mean_yr:n_taxa) %>% 
  # filter(metric != "n_taxa") %>% 
  mutate(
    reverse = abs(value-2100)  + 2000
  )

# Estimate the value (catch and USD) of emerging species
Eme_value_data <- Eme_Res %>% 
  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key) %>% 
  summarise(n()) %>% 
  left_join(Sau_data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  filter(!is.na(variable)) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
  summarise(
    sum_value = sum(value,na.rm=T),
    n_taxa = length(unique(taxon_key)) 
  )

# Total value of each fishing nation

Total_value <- Sau_data %>% 
  group_by(fishing_entity_id,variable) %>% 
  summarise(tot_value= sum(value,na.rm=T))

Value_map_data <- Eme_value_data %>% 
  left_join(Total_value,
            by = c("fishing_entity_id","variable")) %>% 
  mutate(participation = (sum_value/tot_value)*100+2000)

# Per fishing entity

fish_d_hist <- Eme_Res %>% 
  group_by(sau_fishing_entity,tresh) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(
    value = length(unique(taxon_key))
  ) %>% 
  filter(!is.na(sau_fishing_entity))

World_map_nemer <- World_map %>%
  full_join(Value_map_data,
            by = c("sau_fishing_entity","fishing_entity_id")
  ) %>% 
  filter(!is.na(participation)) %>% 
  sf::st_transform(crs = "+proj=eck4")

# These are regions without transboundary species
grey_land <- World_map %>% 
  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
  filter(!sovereignt %in% c("Antarctica"))%>% 
  sf::st_transform(crs = "+proj=eck4")

```

###### Plot ToE map

```{r emergence_map_plot, eval = F}
# -------------------------- #
# Call plot for both tresholds
# -------------------------- #
tresh_list <- c("tresh_one","tresh_two")
for(i in 1:2){

  name_save <- paste("./Figures/emer_map_lat_",tresh_list[i],".png",sep = "")

  # Value Map
  World_map_part <- World_map_nemer %>% 
    filter(
      tresh ==  tresh_list[i],
      variable == "value"
    )

  # ToE Map 
  toe_map_data <- World_eez_sf %>% 
    left_join(Eez_d_hist,
              by = "eezid") %>% 
    filter(tresh == tresh_list[i],
           metric == "mean_yr")%>% 
    sf::st_transform(crs = "+proj=eck4")

  # Map with countries without ToE

  russia_churchil <- World_eez_sf %>% 
    filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))%>% 
    sf::st_transform(crs = "+proj=eck4")

  no_toe_eez <- World_eez_sf %>% 
    filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
           !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
    rbind(russia_churchil)%>% 
    sf::st_transform(crs = "+proj=eck4")

  no_toe_land <- World_map %>% 
    filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
           !name %in% World_map_part$name)%>% 
    sf::st_transform(crs = "+proj=eck4")  # Filter EEZs with ToE stocks

  map <- ggplot(toe_map_data) +
    geom_sf(
      aes(
        fill = reverse #original map
        # fill = value
      ),
      color =  "white",
      size = 0.1
    ) +
    # Inlcude the participation of species
    geom_sf(data = World_map_part,
            # aes(fill = participation+2000), #original map
            aes(fill = participation, 
                color = participation),
            # color =  "white",
            size = 0.1
    ) +
    # grey maps
    # ggplot() +
    geom_sf(data = grey_land, aes(), fill = "grey50", 
            # color =  "white",
            size = 0.1) +
    # geom_sf(data = grey_eez, aes(), fill = "white", color =  "grey50",size = 0.1) +
    geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
    # Set Participation labels
    annotate("text",
             x = -14012580,
             y = -8357450,
             label = "Transboundary fish revenue (%)",
             color="black",
             size=5,
             angle=0,
             fontface="bold",
             hjust=0) +
    annotate("text",
             x = c(-3812580,-4500,3800000,7700000,11512580),
             y = -8400000,
             # x =(c(-8500000,-3000000,2500000,7500000,12800000)),
             # y = -9050000,
             label = c(0,25,50,75,100),
             color="black",
             size=4,
             angle=0,
             # fontface="bold",
             hjust=0) +
    # # Set ToE labels
    annotate("text",
             # x =-150,
             # y = -85,
             x = -13012580,
             y = -9657450,
             label = "Average year of range shift",
             color="black",
             size=5,
             angle=0,
             fontface="bold",
             hjust=0) +
    annotate("text",
             # x = c(-79,-39,2,42,81),
             # y = -85,
             x = c(-4412580,-430000,3410000,7300000,11312580),
             y = -9657450,
             # label = seq(2100,2000,-25), # original
             label = c(">2100",seq(2075,2000,-25)), # not inluding the > because we dont know
             color="black",
             size=4,
             angle=0,
             # fontface="bold",
             hjust=0) +
    # # Set No emergence label
    annotate("text",
             # x =110,
             # y = -77,
             x = 13000000,
             y = -9000000,
             label = "No shift",
             color="black",
             size=5,
             angle=0,
             fontface="bold",
             hjust=0) +
    annotate("rect", 
             xmin = 15500000, 
             xmax = 16500000, 
             ymin = -8600000, 
             ymax = -9300000,
             fill = "lightblue") +
    scale_fill_viridis("",
                       limits = c(2000,2100),
                       breaks = seq(2000,2100,25),
                       labels = c("","","","","")#,
                       # direction = -1
    ) +
    scale_color_viridis("",
                       limits = c(2000,2100),
                       breaks = seq(2000,2100,25),
                       labels = c("","","","","")#,
                       # direction = -1
    ) +
    my_ggtheme_m(leg_pos = "bottom") +
    theme(
      legend.box.margin = margin(t = -61,l = 180),
      legend.text = element_blank(),
      legend.background = element_blank(),
      legend.key.width = unit(5, "line")
    )

  ## --------------- ##
  # Regional Boxplot
  ## --------------- ##

  # Weighted verion
  # boxplot_data <- Eme_Res %>% 
  #   # Get the order of boxes
  #   filter(tresh == tresh_list[i],
  #          !is.na(un_sub_region)) %>%
  #   group_by(un_sub_region) %>%
  #   summarise(n_eezs = length(unique(sf_eez_name))) %>%
  #   # Get the rest of the data
  #   left_join(Eme_Res) %>%
  #   filter(tresh == tresh_list[i],
  #          !is.na(un_sub_region)) %>%
  #   # mutate(weight = )
  #   mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
  #                                 ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
  #                                        ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
  #                                               paste(un_sub_region)))
  #   )#,
  #   # For number labels
  #   # un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
  #   )

  # Non-weighted verion

  boxplot_data <- Eme_Res %>%
    # Get the order of boxes
    filter(tresh == tresh_list[i],
           !is.na(un_sub_region)) %>%
    group_by(un_sub_region) %>%
    summarise(order = median(emerging_yr),
              n_eezs = length(unique(sf_eez_name))) %>%
    # Get the rest of the data
    left_join(Eme_Res) %>%
    filter(tresh == tresh_list[i],
           !is.na(un_sub_region)) %>%
    mutate(un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and\nthe Car.",
                                  ifelse(str_detect(un_sub_region,"Australia"),"Aus. and New Z.",
                                         ifelse(str_detect(un_sub_region,"Sub-Saharan"),"Sub-Saharan\nAfrica",
                                                paste(un_sub_region)))
    ),
    # For number labels
    un_sub_region = paste(un_sub_region," (n = ",n_eezs,")",sep="")
    )

  boxplot_data$un_sub_region <- boxplot_data$un_sub_region %>% 
    gsub("Northern","N.",.) %>% 
    gsub("Southern","S.",.) %>% 
    gsub("Western","W.",.) %>% 
    gsub("Eastern","E.",.) %>% 
    gsub("South-eastern","S.E.",.)

  # ---------- #
  # Summary data
  # ---------- #

  medians_data <- boxplot_data %>% 
    group_by(un_sub_region,un_geo_region) %>% 
    summarise(mean = mean(emerging_yr,na.rm = T),
              sd = sd(emerging_yr,na.rm = T),
              min = min(emerging_yr,na.rm = T),
              max = max(emerging_yr,na.rm = T),
              med = median(emerging_yr,na.rm = T),
              n_stocks = n()
    )

  boxplot_data <- boxplot_data %>% 
    left_join(medians_data,
              by = c("un_sub_region","un_geo_region")
    )

  boxplot <- ggplot(boxplot_data) +
    geom_violin(
      aes(
        x = reorder(un_sub_region,order),
        y = emerging_yr
      ),
      fill = "grey"
    ) + 
    facet_wrap(~un_geo_region,
               scales = "free_x",
               nrow = 1) +
    # Median points
    geom_point(
      aes(
        x = reorder(un_sub_region,order),
        y = mean,
        colour = mean
      ),
      size = 5
    ) + 
    # h line
    geom_hline(yintercept = 2030, color = "grey", size = 0.5, linetype="dashed") +
    scale_colour_viridis("",
                         limits = c(2000,2100),
                         breaks = seq(2000,2100,25),
                         labels = c("","","","",""),
                         direction = -1
    ) +
    xlab("") +
    ylab("Year of range shift") +
    scale_y_continuous(
      labels = c(2000,2025, 2050, 2075, ">2100"), # not inluding the > because we dont know
    ) +
    my_ggtheme_p() +
    theme(
      axis.text.x = element_text(size = 16, angle = 45, face = "plain", hjust = 1), 
      strip.text.x = element_text(size = 18, colour = "black"), 
      axis.title.y = element_text(size = 18),
      axis.text.y = element_text(size = 16), 
      panel.background = element_rect(colour = "white", fill = "white"),
      # axis.ticks.x = element_blank(),
      legend.position = ""
    );boxplot

  ## --------------- ##
  # Combine plots and save them
  ## --------------- ##

  p <- ggdraw() +
    # ToE Latitudinal plot
    # draw_plot(Eez_hist, x = 0.873, y = 0.53, width = 0.124, height = 0.37) +
    # Time of Emergence Map
    draw_plot(map, x = 0, y = 0.39, width = 1, height = 0.70) +
    # Regional boxplot
    draw_plot(boxplot, x = 0, y = -0.03, width = 1, height = 0.44) +
    # Labels
    draw_plot_label(label = c("A", "B"), 
                    size = 20,
                    x = c(0, 0), 
                    y = c(0.98, 0.46)
    ) + 
    theme(plot.background = element_rect(fill="white", color = NA))

  save_plot(
    name_save,
    # "./Figures/test.png",
    p,
    base_height = 11,
    base_width = 12
  )
}

```

###### Splitted Figure 3 for supplements

Note this chunk needs `emergence_map_plot` to be ran.

Year of change in the shared distribution of 4,119 transboundary stocks and their contribution to fishing revenue. A) Contribution of shifting stocks to a country or territory's total fishing revenue from transboundary stocks. B) Average year of range shift across an EEZ’s transboundary stocks. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue. This figure is a detached version of Figure 3

```{r supplement_figure_splitted}

# Value Map
World_map_part <- World_map_nemer %>% 
  filter(
    tresh ==  "tresh_two",
    variable == "value"
  )

# ToE Map 
toe_map_data <- World_eez_sf %>% 
  left_join(Eez_d_hist,
            by = "eezid") %>% 
  filter(tresh == "tresh_two",
         metric == "mean_yr")%>% 
  sf::st_transform(crs = "+proj=eck4")

# Map with countries without ToE

russia_churchil <- World_eez_sf %>% 
  filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))%>% 
  sf::st_transform(crs = "+proj=eck4")

no_toe_eez <- World_eez_sf %>% 
  filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
         !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
  rbind(russia_churchil)%>% 
  sf::st_transform(crs = "+proj=eck4") %>% 
  mutate(value = NA)

no_toe_land <- World_map %>% 
  filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
         !name %in% World_map_part$name)%>% 
  sf::st_transform(crs = "+proj=eck4") %>%   # Filter EEZs with ToE stocks 
  mutate(value = NA)

# EEZ map
eez_map <- ggplot(toe_map_data) +
  geom_sf(aes(fill = value), # for 2000 to 2100),
          color =  "white",
          size = 0.1
  ) +
  geom_sf(data = grey_land, aes(),
          fill = "grey50", 
          color =  "white",
          size = 0.1) +
  geom_sf(data = no_toe_eez, aes(fill = as.numeric(value)),
          color =  "white",
          size = 0.1) +
  scale_fill_viridis("Average year of range shift",
                     limits = c(2000,2100),
                     breaks = seq(2000,2100,25),
                     labels = seq(2000,2100,25),
                     direction = -1,
                     na.value = "lightblue"
  ) +
  my_ggtheme_m(leg_pos = "bottom") +
  theme(legend.key.width = unit(3, "line")) +
  guides(
    fill=guide_colorbar(title.position="top", 
                        title.hjust =0.5)
  )

# Land Map
land_map <- ggplot() +
  geom_sf(data = World_map_part,
          aes(fill = participation-2000, 
              color = participation-2000),
          size = 0.1
  )  +
  geom_sf(data = no_toe_land, aes(fill = as.numeric(value),
                                  color = as.numeric(value)),
          size = 0.1) +
  scale_fill_viridis("Transboundary fish revenue (%)",
                     limits = c(0,100),
                     breaks = seq(0,100,25),
                     labels = seq(0,100,25),
                     direction = 1,
                     na.value = "lightblue",
                     option = "A"
  ) +
  scale_color_viridis("Transboundary fish revenue (%)",
                      limits = c(0,100),
                      breaks = seq(0,100,25),
                      labels = seq(0,100,25),
                      direction = 1,
                      na.value = "lightblue",
                      option = "A"
  ) +
  my_ggtheme_m(leg_pos = "bottom") +
  theme(legend.key.width = unit(3, "line")) +
  guides(
    fill=guide_colorbar(title.position="top",
                        title.hjust =0.5)
  )

 ## --------------- ##
# Combine plots and save them
## --------------- ##

p <- ggdraw() +
  draw_plot(eez_map, x = 0, y = 0, width = 1, height = 0.5) +
  # Regional boxplot
  draw_plot(land_map, x = 0, y = 0.5, width = 1, height = 0.5) +
  # Labels
  draw_plot_label(label = c("A", "B"), 
                  size = 20,
                  x = c(0, 0), 
                  y = c(0.98, 0.46)
  ) + 
  theme(plot.background = element_rect(fill="white", color = NA))

save_plot(
  "./Figures/FigSplitSX.png",
  p,
  base_height = 10,
  base_width = 10
)

```

##### Fig. 4 Mapping the time of emergence of Top stocks

```{r emergence_map, eval = F}

top_stocks_per_country <- Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
  summarise(n()) %>% 
  left_join(Sau_data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  filter(!is.na(variable)) %>% 
  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
  # Select top 5 most valuable stocks
  top_n(5,value)

# Per EEZs
top_toe <- Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
  summarise_at(vars(mean_yr),
               c("mean"=mean,
                 "sd" = sd)) %>% 
  mutate(
    geo_region = ifelse(is.na(geo_region),"Other",geo_region)
  )

# Numbers for paper
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  group_by(variable,tresh) %>% 
  summarise_at(vars(mean_yr),
               c("mean"=mean,
                 "sd" = sd))

# Number of coastal nations
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  filter(mean_yr <= 2021) %>% 
  group_by(variable,tresh) %>% 
  summarise(
    n_countries = length(unique(sau_fishing_entity))
  ) %>% 
  mutate(per = n_countries/length(unique(Sau_data$fishing_entity_id)))

# Number of shifting stocks per geo-region
Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(eezid,sau_fishing_entity,geo_region,tresh,taxon_key) %>% 
  summarise(mean_yr = mean(emerging_yr,na.rm=T)) %>% 
  right_join(top_stocks_per_country,
             by = c("sau_fishing_entity","geo_region","tresh","taxon_key")
  ) %>% 
  filter(mean_yr <= 2021) %>% 
  group_by(geo_region) %>% 
  tally() %>% 
  arrange(n) %>% 
  top_frac(0.25,n) # top 25%

```

###### The Map option with tops

```{r the map}

# Create data
Top_map_data <- top_toe %>% 
  filter(tresh == selected_tresh,
         variable == "value") # tried with catch but is quite the same

World_map_top <- World_eez_sf %>% 
  left_join(Matching_names) %>% 
  full_join(Top_map_data,
            by = c("sau_fishing_entity")
  ) %>% 
  filter(!is.na(variable)) %>% 
  sf::st_transform(crs = "+proj=eck4")

# Map with countries without ToE

russia_churchil <- World_eez_sf %>% 
  filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))%>% 
  sf::st_transform(crs = "+proj=eck4")

no_toe_eez <- World_eez_sf %>% 
  sf::st_transform(crs = "+proj=eck4") %>% 
  filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
         !eez_name %in% World_map_top$eez_name) %>%  # Filter EEZs with ToE stocks
  rbind(russia_churchil)%>% 
  sf::st_transform(crs = "+proj=eck4")

no_toe_land <- World_map %>% 
  filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
         !name %in% World_map_top$name
  ) %>% 
  sf::st_transform(crs = "+proj=eck4")

map <- ggplot(World_map_top) +
  geom_sf(
    aes(
      fill = mean #original map
      # fill = value
    ),
    color =  "white",
    size = 0.1
  ) +
  # grey maps
  # ggplot() +
  geom_sf(data = subset(World_map, sovereignt != "Antarctica"), aes(), fill = "grey50", color =  "white",size = 0.1) +
  geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  # geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  # Set Participation labels
  scale_fill_viridis("Average year of range shift of top valued transboundary stocks",
                     limits = c(2000,2100),
                     breaks = seq(2000,2100,25),
                     direction = -1
  ) +
  my_ggtheme_m(leg_pos = "bottom") +
  theme(legend.key.width = unit(5, "line")) +
  guides(
    fill=guide_colorbar(title.position="top", 
                        title.hjust =0.5)
  )

ggsave(
  plot = map,
  "./Figures/top_toe_map.png",
  width = 8,
  height = 5,
  units = "in"
)

```

#### Species level analysis

Supplemental box picture by habitat preference

```{r deme_pela, eval = F, echo =F}

Dem_pel <- Eme_Res %>% 
  filter(tresh == selected_tresh,
         !is.na(un_sub_region)) %>% 
  left_join(Spp_env_pref) %>% 
  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",paste(DemersPelag)),
  un_sub_region = ifelse(str_detect(un_sub_region,"Latin"),"Ltn. Ame. and the Car.",paste(un_sub_region)))

# Dem Pelagic boxplot
Eme_Res %>% 
  left_join(Spp_env_pref) %>% 
  mutate(DemersPelag = ifelse(is.na(DemersPelag),"other",DemersPelag)) %>% 
  group_by(DemersPelag,tresh) %>%
  summarise(mean_toe = median(emerging_yr,na.rm=T),
            n_spp = length(unique(taxon_key)),
            n_stock = n()
  ) %>%
  left_join(Dem_pel) %>%
  mutate(
    label = paste0(DemersPelag, "\n(n=",n_stock,")")
  ) %>%
  filter(tresh == selected_tresh,
         DemersPelag != "pelagic") %>% # just one spp.
  ggplot() +
  # geom_violin(
  geom_boxplot(
    aes(x = reorder(label,mean_toe),
        # aes(x = DemersPelag,
        y = emerging_yr,
        fill = mean_toe
    ),
    alpha = 0.8,
    # fill = "grey",
    coef = 0, # to have dotted error bars
    outlier.shape = NA # to have dotted error bars
  ) +
  # to have dotted error bars
  stat_boxplot(
    aes(x = reorder(label,mean_toe),
        # aes(x = DemersPelag,
        y = emerging_yr,
        # fill = DemersPelag
    ),
    geom ='errorbar',
    width = 0.3,
    linetype = "dashed") + 
  geom_hline(yintercept = 2030, color = "grey", size = 0.5, linetype="dashed") +
  xlab("") +
  ylab("Year of range shift") +
  my_ggtheme_p(leg_pos = "right", leg_tl_s =11, leg_tx_s = 10, axx_tx_ang = 45, ax_tx_s = 10, hjust = 1) +
  scale_fill_viridis("Average year of\nrange shift",
                     limits = c(2000,2100),
                     breaks = seq(2000,2100,25),
                     # labels = c("","","","",""),
                     direction = -1
  ) +
  scale_x_reordered() +
  ggsave("./Figures/regional_hab_boxplot.png",
         width = 10,
         height = 5,
         units = "in"
  )

```

### Statistics

#### Test for normallity

```{r, normality_test, eval = F, echo = F}
# Test for normality

Stats_data <- Eme_Res %>% 
  filter(tresh == selected_tresh,
         n_ens == 10)

# ------------------- #
# Test normal assumptions 
# ------------------- #

hist(Stats_data$emerging_yr) # Positively skewed
qqnorm(Stats_data$emerging_yr) # not normally distributed
skewness(Stats_data$emerging_yr) #1.02
kurtosis(Stats_data$emerging_yr) #2.82

# ------------------- #

# (NOT NORMAL DISTRIRBUTION)

```

#### Regional differences

```{r multiple_reg, eval = F, echo = F}

Reg_diff_data <- Eme_Res %>% 
  filter(tresh == selected_tresh,
         !is.na(un_sub_region)
  )

# ------------------- #
# Run non-parametric Krustal
# ------------------- #

Reg_kw <- kruskal.test(un_sub_region~emerging_yr, data = Reg_diff_data)

Reg_kw
# Kruskal-Wallis rank sum test
# 
# data:  un_sub_region by emerging_yr
# Kruskal-Wallis chi-squared = 287.23, df = 93, p-value < 2.2e-16

kruskalmc(emerging_yr~un_sub_region, data = Reg_diff_data) %>% 
  as.data.frame() %>% 
  rownames_to_column("comparrison") %>% 
  # head()
  View()
write_csv("kw_regional.csv")

# Multiple differences but Eastern Asia, Lat. Ame. Car. and Polynesia are the ones who differ the most (e.g. differ with more regions).

# ------------------- #
# Just to double check ANOVA results (give the same)
# ------------------- #
Reg_aov <- aov(emerging_yr~un_sub_region, data = Reg_diff_data)
Reg_aov

summary(Reg_aov)

TukeyHSD(Reg_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) %>%
  View()

```

#### Habitat preference differences

```{r multiple_reg, eval = F, echo = F}

Hab_diff_data <- Eme_Res %>% 
  filter(tresh == selected_tresh,
         !is.na(un_sub_region)
  ) %>% 
  left_join(Spp_env_pref,
            by = "taxon_key")

# ------------------- #
# Run non-parametric Krustal
# ------------------- #

Hab_kw <- kruskal.test(DemersPelag~emerging_yr, data = Hab_diff_data)

Hab_kw
# Kruskal-Wallis rank sum test

# data:  DemersPelag by emerging_yr
# Kruskal-Wallis chi-squared = 286.48, df = 93, p-value < 2.2e-16

kruskalmc(emerging_yr~DemersPelag, data = Hab_diff_data) %>% 
  as.data.frame() %>% 
  rownames_to_column("comparrison") %>% 
  # head()
  View()
write_csv("kw_habitat.csv")

# Multiple differences but main differences for pelagic-ocenaic and reefs

# ------------------- #
# Just to double check ANOVA results (give the same)
# ------------------- #

Hab_aov <- aov(emerging_yr~DemersPelag, data = hab_diff_data)
Hab_aov

summary(Hab_aov)

TukeyHSD(Hab_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) %>%
  View()

```

##  Stock Share Ratio Analysis

### Fig. 5 and S5 Proportion + Netowrk

Note that this map is based on the data viz site: https://www.data-to-viz.com/story/MapConnection.html

#### Prepare data

```{r ssr_map_data, eval = T, echo = T}
# A function that makes a dateframe per connection (we will use these connections to plot each lines)
data_for_connection=function( dep_lon, dep_lat, arr_lon, arr_lat, group){
  inter <- gcIntermediate(c(dep_lon, dep_lat), c(arr_lon, arr_lat), n=50, addStartEnd=TRUE, breakAtDateLine=F)             
  inter=data.frame(inter)
  inter$group=NA
  diff_of_lon=abs(dep_lon) + abs(arr_lon)
  if(diff_of_lon > 180){
    inter$group[ which(inter$lon>=0)]=paste(group, "A",sep="")
    inter$group[ which(inter$lon<0)]=paste(group, "B",sep="")
  }else{
    inter$group=group
  }
  return(inter)
}

# Use a different palet between SSR and ToE
pal <- wesanderson::wes_palette("Zissou1",100,type = "continuous")

# Prepare data from results
Data <- Prop_results %>% 
  # get coords for sources
  left_join(Eez_centroid,
            by = "eez_name") %>%
  rename(source = eez_name,
         start_lon = longitude,
         start_lat = latitude,
         eez_name=eez_neighbour) %>% 
  # get coords for targets
  left_join(Eez_centroid,
            by = "eez_name") %>%
  rename(target = eez_name,
         end_lon = longitude,
         end_lat = latitude) %>% 
  ungroup() %>% 
  # Plot miscs
  mutate(
    per_early = ifelse(is.na(per_early),0,per_early),
    per_mid = ifelse(is.na(per_mid),0,per_mid)
  ) %>% 
  drop_na()

# Get paths data
path_data <- Data %>% 
  filter(
    variable %in% c("mean_eez")
  )

data_ready_plot=data.frame()

for(i in c(1:nrow(path_data))){
  tmp=data_for_connection(path_data$start_lon[i], path_data$start_lat[i], path_data$end_lon[i], path_data$end_lat[i] , i)
  tmp$per_early_plot=path_data$per_early_plot[i]
  tmp$per_mid_plot=path_data$per_mid_plot[i]
  tmp$source=path_data$source[i]
  tmp$objective=path_data$target[i]
  data_ready_plot=rbind(data_ready_plot,tmp)
}

data_ready_plot <- data_ready_plot %>% 
  mutate(
    per_early = ifelse(per_early_plot >= 50,50,per_early_plot),
    per_mid = ifelse(per_mid_plot >= 50,50,per_mid_plot)) %>% 
  gather("time_step","percentage",per_early,per_mid)

```

#### Select a time step

```{r}
# SELECT A TIME STEP
# time_toplot <- "per_mid"
time_toplot <- "per_early"
```

#### Get data ready for time step

```{r ssr_map_data, eval = T, echo = T}

sf_ready_plot <- data_ready_plot %>% 
  filter(time_step == time_toplot) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  st_set_crs(4326) %>% # to match shapefiles
  st_transform(crs = "+proj=eck4")

# Get arrow data
arrows_data <- as.data.frame(st_coordinates(st_cast(sf_ready_plot,"POINT")$geometry)) %>% 
  bind_cols(data_ready_plot %>% 
              filter(time_step == time_toplot)
  )

# Get point data
point_data <- path_data %>% 
  gather("time_step","percentage",per_early:per_mid) %>% 
  filter(time_step == time_toplot,
         percentage <= 0) %>% 
  st_as_sf(
    coords = c("start_lon", "start_lat")
  ) %>% 
  select(source,target,geometry) %>% 
  st_set_crs(4326) %>% 
  st_transform(crs = "+proj=eck4")

#### Get the change of species

n_trans_eez <- Transboundary_spp %>% 
  group_by(eez_name) %>% 
  summarise(
    # n_trans = unique(length(taxon_key))
    n_trans = n()
  )

# Estimate the percentage changing
n_taxa_data <- Prop_results %>% 
  filter(variable == "n_spp") %>%
  select(eez_name,early,mid) %>% 
  left_join(n_trans_eez) %>% 
  # mutate(per = (early/n_trans)*100) %>% 
  mutate_at(vars(early,mid),
            .funs = ~(./n_trans)*100) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(Matching_names) %>% 
  select(sau_fishing_entity,fishing_entity_id,early,mid) %>% 
  gather("time_step","avrg_spp_per",early:mid) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,time_step) %>%
  summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>% 
  mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>%
  mutate(avrg_spp_per_100 = avrg_spp_per,
         avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per)) %>% 
  filter(time_step == time_toplot)

# Included in map
n_taxa_map <- World_map %>% 
  full_join(n_taxa_data,
            by = c("sau_fishing_entity","fishing_entity_id")
  )

# Grey land
grey_land <- World_map %>% 
  filter(is.na(fishing_entity_id)) %>% 
  filter(!sovereignt %in% c("Antarctica"))
```

#### Plot SSR map

```{r ssr_map_plot, eval = T, echo = T}
# Plot
p <-
  ggplot() +
  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
          aes(fill = avrg_spp_per),
          size = 0.1,
          col = "white") +
  geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
  geom_sf(data = point_data, aes(), color = "grey90") +
  geom_path(data = subset(arrows_data, percentage >= 0),
            aes(X, Y,
                group = group,
                col = percentage
            ),
            size = 1,
            arrow = arrow(length = unit(0.01, "npc"),
                          ends = "first"),
            show.legend = NA) +
  my_ggtheme_m(leg_pos = "bottom") +
  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                        colours = pal,
                        limits = c(0,max(sf_ready_plot$percentage,na.rm=T)),
                        breaks = seq(0,max(sf_ready_plot$percentage,na.rm=T),10),
                        labels = c(seq(0,max(sf_ready_plot$percentage-10,na.rm=T),10),">50")
  ) +
  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                       colours = pal,
                       limits = c(0,max(sf_ready_plot$percentage,na.rm=T)),
                       breaks = seq(0,max(sf_ready_plot$percentage,na.rm=T),10),
                       labels = c(seq(0,max(sf_ready_plot$percentage-10,na.rm=T),10),">50")
  ) + guides(
    color=guide_colorbar(title.position="top")
  ) +
  theme(legend.key.width = unit(4,"line"))

# name <- paste0("./Figures/ssr_",time_toplot,".png")
# Save figure
# ggsave(plot =  p,
#        name,
#        width = 8,
#        height = 5,
#        units = "in")

```

#### Regional maps for supplements

Note that the following chunk needs `ssr_map_data` to work.

```{r}

## ------------------- ##
# Latin america Map
## ------------------- ##

lat_am_map <- ggplot() +
  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
          aes(fill = avrg_spp_per),
          size = 0.1,
          col = "white") +
  # geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
  geom_sf(data = point_data, aes(), color = "grey90") +
  geom_path(data = subset(arrows_data, percentage >= 0),
            aes(X, Y,
                group = group,
                col = percentage
            ),
            size = 1,
            arrow = arrow(length = unit(0.03, "npc"),
                          ends = "first"),
            show.legend = NA) +
  coord_sf(xlim = c(-9000000,-5000000),ylim = c(1000000, 4000000)) + 
  labs(x = "", y = "") +
  # scale_x_continuous(breaks = seq(-100,60,10)) +
  # scale_y_continuous(breaks = seq(10,30,10)) +
  my_ggtheme_m(leg_pos = "bottom") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"),
    panel.grid.major =  element_line(color = "white")
  ) +
  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                        colours = pal,
                        limits = c(0,max(Lat_am_data$percentage)),
                        breaks = seq(0,max(Lat_am_data$percentage),10)
  ) +
  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                       colours = pal,
                       limits = c(0,max(Lat_am_data$percentage)),
                       breaks = seq(0,max(Lat_am_data$percentage),10)
  ) + guides(
    color=guide_colorbar(title.position="top", 
                         title.hjust =0.5)
  ) +
  theme(legend.key.width = unit(4,"line"))

ggsave(filename = "lat_am_mid.png",
       lat_am_map,
       width = 6,
       height = 6)

## ------------------- ##
# Northern Europe Map
## ------------------- ##

EU_map <- ggplot() +
  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
          aes(fill = avrg_spp_per),
          size = 0.1,
          col = "white") +
  geom_sf(data = subset(sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
  geom_sf(data = point_data, aes(), color = "grey90") +
  geom_path(data = subset(arrows_data, percentage >= 0),
            aes(X, Y,
                group = group,
                col = percentage
            ),
            size = 1,
            arrow = arrow(length = unit(0.03, "npc"),
                          ends = "first"),
            show.legend = NA) +
  coord_sf(xlim = c(-1300000,2000000),ylim = c(5500000, 7500000))  +
  labs(x = "", y = "") +
  scale_x_continuous(breaks = seq(-100,60,10)) +
  scale_y_continuous(breaks = seq(10,30,10)) +
  my_ggtheme_m(leg_pos = "bottom") +
  theme(
    panel.background = element_rect(fill = "white", color = "white"),
    panel.grid.major =  element_line(color = "white")
  ) +
  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                        colours = pal,
                        limits = c(0,max(EU_data$percentage)),
                        breaks = seq(0,max(EU_data$percentage),10)
  ) +
  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                       colours = pal,
                       limits = c(0,max(EU_data$percentage)),
                       breaks = seq(0,max(EU_data$percentage),10)
  ) + guides(
    color=guide_colorbar(title.position="top",
                         title.hjust =0.5)
  ) +
  theme(legend.key.width = unit(4,"line"))

ggsave(filename = "EU_early.png",
       EU_map,
       width = 6,
       height = 6)

```

#### SSR TOP 5 figure

```{r fig_map_Juan, eval = F, echo =F}

top_stocks_per_country <- Eme_Res %>% 
  rename(geo_region = un_sub_region) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,geo_region,tresh,taxon_key) %>% 
  summarise(n()) %>% 
  left_join(Sau_data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  filter(!is.na(variable)) %>% 
  # group_by(sau_fishing_entity,fishing_entity_id,tresh,variable) %>% 
  # summarise(sum_value = sum(value,na.rm=T)) %>% 
  group_by(sau_fishing_entity,geo_region,tresh,variable) %>% 
  # Select top 5 most valuable stocks
  top_n(5,value)

top_taxon_list <-  unique(top_stocks_per_country$taxon_key)
# Read proportion results
suppressMessages(
  top_prop_results <- PropChange(top_taxon_list, links="y",top = T) %>%
    filter(!is.na(eez_neighbour)) %>%
    mutate(per_mid_plot =ifelse(per_mid >= 100, 100,
                                ifelse(per_mid <= -100,-100,per_mid)
    ),
    per_early_plot =ifelse(per_early >= 100, 100,
                           ifelse(per_early <= -100,-100,per_early)
    )
    )
)

### Median number of stocks changing in each time step

top_prop_results %>% 
  filter(variable == "n_per") %>% 
  group_by() %>% 
  summarise_at(vars(early,mid),funs(
    median(.,na.rm=T),
    sd(.,na.rm=T),
    length(unique(eez_name)),
    length(unique(eez_name))/280*100
  )
  )

# Prepare data from results
top_data <- top_prop_results %>% 
  # get coords for sources
  left_join(Eez_centroid,
            by = "eez_name") %>%
  rename(source = eez_name,
         start_lon = longitude,
         start_lat = latitude,
         eez_name=eez_neighbour) %>% 
  # get coords for targets
  left_join(Eez_centroid,
            by = "eez_name") %>% 
  rename(target = eez_name,
         end_lon = longitude,
         end_lat = latitude) %>% 
  ungroup() %>% 
  # Plot miscs
  mutate(
    per_early = ifelse(is.na(per_early),0,per_early),
    per_mid = ifelse(is.na(per_mid),0,per_mid)
  ) %>% 
  drop_na() %>%
  # gather("variable","value",per_early:per_early_plot) %>% 
  filter(variable %in% c("mean_eez"))

top_data_ready_plot=data.frame()

for(i in c(1:nrow(top_data))){
  tmp=data_for_connection(top_data$start_lon[i], top_data$start_lat[i], top_data$end_lon[i], top_data$end_lat[i] , i)
  tmp$per_early_plot=top_data$per_early_plot[i]
  tmp$per_mid_plot=top_data$per_mid_plot[i]
  tmp$source=top_data$source[i]
  tmp$objective=top_data$target[i]
  top_data_ready_plot=rbind(top_data_ready_plot,tmp)
}

top_data_ready_plot <- top_data_ready_plot %>% 
  mutate(
    per_early = ifelse(per_early_plot >= 50,50,per_early_plot),
    per_mid = ifelse(per_mid_plot >= 50,50,per_mid_plot)) %>% 
  gather("time_step","percentage",per_early,per_mid)

top_sf_ready_plot <- top_data_ready_plot %>% 
  filter(time_step == "per_early") %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  st_set_crs(4326) %>% # to match shapefiles
  st_transform(crs = "+proj=eck4")

# Get arrow data
top_arrows_data <- as.data.frame(st_coordinates(st_cast(top_sf_ready_plot,"POINT")$geometry)) %>% 
  bind_cols(top_data_ready_plot %>% 
              filter(time_step == "per_early")
  )

# Get point data
top_point_data <- top_data %>% 
  gather("time_step","percentage",per_early:per_mid) %>% 
  filter(time_step == "per_early",
         percentage <= 0) %>% 
  st_as_sf(
    coords = c("start_lon", "start_lat")
  ) %>% 
  select(source,target,geometry) %>% 
  st_set_crs(4326) %>% 
  st_transform(crs = "+proj=eck4")

# # # Estimate the percentage changing
# n_taxa_data <- top_prop_results %>%
#   filter(variable == "n_spp") %>%
#   select(eez_name,early,mid) %>%
#   left_join(n_trans_eez) %>%
#   # mutate(per = (early/n_trans)*100) %>%
#   mutate_at(vars(early,mid),
#             .funs = ~(./n_trans)*100) %>%
#   rename(sf_eez_name = eez_name) %>%
#   left_join(Matching_names) %>%
#   select(sau_fishing_entity,fishing_entity_id,early,mid) %>%
#   gather("time_step","avrg_spp_per",early:mid) %>%
#   group_by(sau_fishing_entity,fishing_entity_id,time_step) %>%
#   summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>%
#   mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>%
#   mutate(avrg_spp_per_100 = avrg_spp_per,
#          avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per)) %>%
#   filter(time_step == time_toplot)

# Included in map
# n_taxa_map <- World_map %>% 
#   full_join(n_taxa_data,
#             by = c("sau_fishing_entity","fishing_entity_id")
#   )

```

```{r}
top_ssr_map <-
  ggplot() +
  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
  # geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica"),
  #         aes(fill = avrg_spp_per),
  #         size = 0.1,
  #         col = "white") +
  geom_sf(data = subset(World_map,sovereignt != "Antarctica"),
          size = 0.1, fill = "grey90", col = "gray30") +
  geom_sf(data = subset(top_sf_ready_plot, percentage >= 0), aes(color = percentage), size = 0.1, fill = "white") +
  geom_sf(data = top_point_data, aes(), color = "grey90") +
  geom_path(data = subset(top_arrows_data, percentage >= 0),
            aes(X, Y,
                group = group,
                col = percentage
            ),
            size = 1,
            arrow = arrow(length = unit(0.01, "npc"),
                          ends = "first"),
            show.legend = NA) +
  my_ggtheme_m(leg_pos = "bottom") +
  scale_color_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                        colours = pal,
                        limits = c(0,max(top_sf_ready_plot$percentage,na.rm=T)),
                        breaks = seq(0,max(top_sf_ready_plot$percentage,na.rm=T),10),
                        labels = c(seq(0,max(top_sf_ready_plot$percentage-10,na.rm=T),10),">50")
  ) +
  scale_fill_gradientn(name= "Changes in catch proportion and changing stocks (%)",
                       colours = pal,
                       limits = c(0,max(top_sf_ready_plot$percentage,na.rm=T)),
                       breaks = seq(0,max(top_sf_ready_plot$percentage,na.rm=T),10),
                       labels = c(seq(0,max(top_sf_ready_plot$percentage-10,na.rm=T),10),">50")
  ) + guides(
    color=guide_colorbar(title.position="top")
  ) +
  theme(legend.key.width = unit(4,"line"))

# Save figure
ggsave(plot =  top_ssr_map,
       "./Figures/ssr_top.png",
       width = 8,
       height = 5,
       units = "in")

```

##### Fig S2 (Histogram of change)

```{r hist_change}

Prop_results %>% 
  gather("time_step","change",per_early,per_mid) %>% 
  filter(variable == "mean_eez" & 
           !is.na(change)
  ) %>%
  ggplot() +
  geom_density(
    aes(
      x = ifelse(change >= 100,100,change),
      # x = change,
      fill = ifelse(time_step == "per_early","q",time_step)
      # fill = time_step,

    ),
    # stat = "count",
    # alpha= 0.8,
    size=0.1,
    # position = "stack"
  ) +
  scale_y_continuous("Density",
                     # breaks = seq(0,90,10),
                     # limits = c(0,100)
  ) +
  scale_x_continuous("Percentage Change",
                     breaks = seq(-75,500,25)
  ) +
  my_ggtheme_p(leg_pos = "right") +
  scale_fill_viridis(discrete = T, 
                     option = "plasma",
                     alpha = 0.5,
                     name = "Time Step", labels = c("Middle", "Early")
  ) +
  coord_flip() +
  ggsave("./Figures/hist_change.png",
         width = 6,
         height = 6,
         units = "in"
  )

```

## Sensitivity analysis

### Testing different Per values for the centroids

```{r data_loading, eval = F, echo = F}

# SAU exploited species list
Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
  clean_names() %>% 
  pull(taxon_name)

# Load all sp in one table
sen_list <- c(20,50,90,95)

sen <- 95
read_sen <- function(sen){

  if(sen != 95){

    level <- paste0("Emergence_2005_p",sen)
    extra_path <- paste0("Toe_Sens_per/",level)

    # Get taxon list
    Taxon_list_sens <- list.files(my_path("R",extra_path = extra_path))

    Paths_sen <- paste(my_path("R",extra_path = extra_path),Taxon_list_sens,sep="")

  }else{

    Taxon_list <- gsub("_.*","",Taxon_list_sens)

    Paths_sen <- paste(my_path("R",extra_path = "Emergence_2005"),Taxon_list_sens,"_emergence.txt",sep="")

  }

  Eme_sen <- bind_rows(lapply(Paths_sen, fread)) %>% 
    mutate(sen = sen) %>% 
    filter(
      n_ens == 10
    ) %>% 
    left_join(Neighbours,
              by = c("eez_name","eez_neighbour")
    ) %>% 
    rename(sf_eez_name=eez_name) %>% 
    left_join(Matching_names,
              by=  "sf_eez_name") %>% 
    clean_names()

  return(Eme_sen)

}

Sen_data <- bind_rows(
  lapply(sen_list,read_sen)
) %>% 
  filter(
    tresh== "tresh_two"
  )

unique(Sen_data$sen)

```

```{r data_analysis, eval = F, echo = F}

# Per Stock
sen_stock_d <-  
  Sen_data %>% 
  left_join(Neighbours) %>%
  group_by(taxon_key,neighbour_id,sen) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,sen) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(sen) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "stock"
  )

#  EEZ data
sen_eez_d <-  Sen_data %>% 
  left_join(Neighbours) %>% 
  filter(n_ens == 10) %>% 
  group_by(sf_eez_name,sen) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,sen) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(sen) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "eez"
  )

# Plot Data
sen_plot_data <- sen_stock_d %>% 
  bind_rows(sen_eez_d) %>% 
  mutate(
    # total = ifelse(level == "stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$sf_eez_name))),
    total = ifelse(level == "stock",max(sen_stock_d$com_sum),max(sen_eez_d$com_sum)),
    per = com_sum/total*100
  ) %>% 
  gather("metric","value",com_sum,per) %>% 
  ungroup() %>% 
  mutate(
    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
    label = ifelse(total == max(sen_stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs"),
    sen = paste("Percentile above",sen,"%") 
  )

# Lines data
dashed_lines <- tibble(
  x = c(2030,2050,2080),
  label = c("Agenda 2030","Early range shift  ","Late range shift   ")
)

# Plot me
ggplot() +
  geom_area(data = subset(sen_plot_data, level == "stock"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  )  +
  geom_area(data = subset(sen_plot_data, level == "eez"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  )  +
  # Shameslly plot it again to fix colors
  geom_area(data = subset(sen_plot_data, level == "stock" & my_facet == "Percentage (%)"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  ) +
  facet_wrap(~my_facet + sen,
             scales = "free",
             nrow=2) + 
  scale_y_continuous("",
                     expand = c(0,0)
  ) +
  scale_x_continuous("Year of range shift",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100"),
                     expand = c(0,0)
  ) +
  # Add dashed lines and labels
  geom_vline(
    data = dashed_lines,
    aes(
      xintercept = x,
      linetype = label,
    ),
    color = "grey"
  ) + 
  my_ggtheme_p() +
  theme(
    legend.position = "right",
    panel.spacing = unit(3, "lines") # space between plots
  ) +
  guides(fill=guide_legend(title= "", reverse=FALSE),
         linetype = guide_legend(title = "Time step")) +
  scale_fill_viridis(discrete = T,
                     begin = 0.1, 
                     end = 0.7,
                     labels = c("EEZs","Stocks")
  ) +
  ggsave("./Figures/Supp3.png",
         width = 14,
         height = 10,
         units = "in"
  )

```

### Noise for all ensemble members

```{r read_raw_data, eval = T, echo = F}

# --------------- #
# Load year of emergence results
# --------------- #

fread2 <- function(x){

  tb <- fread(x)

  if(nrow(tb) >0){
    return(tb)
  }

}

sens_paths <- list.files(my_path("R","Emergence_alt_sens"),full.names = T)

# Load all sp in one table
Eme_Alt_Res <- bind_rows(lapply(sens_paths, fread2)) %>% 
  mutate(method= "alt") %>% 
  left_join(Neighbours,
            by = c("eez_name","eez_neighbour")
  ) %>% 
  rename(sf_eez_name=eez_name) %>% 
  left_join(Matching_names,
            by=  "sf_eez_name") %>% 
  clean_names() %>% 
  mutate(tresh = str_to_lower(tresh)) %>% 
  filter(emerging_yr != 2100)

# Combine both datas
sens_data <- Eme_Res %>% 
  filter(taxon_key %in% Eme_Alt_Res$taxon_key) %>% 
  mutate(method = "org") %>% 
  bind_rows(Eme_Alt_Res)

n_stock <- sens_data %>% 
  filter(method == "org") %>% 
  group_by(neighbour_id) %>% 
  summarise(n_stock = length(unique(taxon_key))) %>% 
  # View()
  group_by() %>% 
  summarise(t_trans_stocks = sum(n_stock)) %>% 
  pull(t_trans_stocks)

n_eez <- length(unique(sens_data$sf_eez_name))

```

#### General numbers

```{r sens_numbers, eval = T, echo = F}

# Total number of emerging stocks

# Global estimates
sens_data %>% 
  group_by(tresh,neighbour_id,method) %>% 
  summarise(
    emerging_stocks = length(unique(taxon_key))
  ) %>% 
  group_by(tresh,method) %>% 
  summarise(
    emerging_stocks = sum(emerging_stocks),
    prop_total = (emerging_stocks/n_stock)*100,
  )

# Looks like there is no difference in the total number of emerging stocks
# tresh     method emerging_stocks prop_total
# 1 tresh_one alt               6665      116. 
# 2 tresh_one org               5729      100  
# 3 tresh_two alt               5258       91.8
# 4 tresh_two org               4108       71.7

# Global average ToE
sens_data %>% 
  group_by(tresh,method) %>% 
  summarise(
    # emerging_stocks = length(unique(taxon_key)),
    # prop_total = (emerging_stocks/9132)*100,
    mean= mean(emerging_yr),
    sd = sd(emerging_yr),
    min = min(emerging_yr),
    max = max(emerging_yr)
  )

# Looks like the alternative method sets a later mean ToE but no real difference in s.d. min nor max
#  tresh     method  mean    sd   min   max
# 1 tresh_one alt    2055.  28.7  2007  2099
# 2 tresh_one org    2029.  26.7  2006  2099
# 3 tresh_two alt    2057.  28.6  2007  2099
# 4 tresh_two org    2036.  28.1  2006  2099

```

#### Area plots 
```{r read_raw_data, eval = T, echo = F}

# Per Stock
stock_d <-  
  sens_data %>% 
  # left_join(Neighbours) %>% 
  # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
  group_by(taxon_key,neighbour_id,tresh,method) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,tresh,method) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(tresh,method) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "stock"
  )

# max(stock_d$com_sum)

#  EEZ data
eez_d <-  sens_data %>% 
  # left_join(Neighbours) %>% 
  filter(n_ens == 10) %>% 
  # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
  group_by(sf_eez_name,tresh,method) %>% 
  summarise(
    taxon_emer = min(emerging_yr)
  ) %>% 
  group_by(taxon_emer,tresh,method) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(tresh,method) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "eez"
  )

# max(eez_d$com_sum)

# Plot Data
plot_data <- stock_d %>% 
  bind_rows(eez_d) %>% 
  filter(tresh != "NA") %>% 
  mutate(
    total = ifelse(level == "stock",n_stock,n_eez),
    per = com_sum/total*100
  ) %>% 
  gather("metric","value",com_sum,per) %>% 
  filter(tresh == "tresh_two")

# 
# # Manually include rest of data until 2100
# cont_eez <- plot_data %>% 
#   filter(level == "eez") %>% 
#   arrange(taxon_emer) %>% 
#   slice(1:24) %>%
#   select(-1) %>% 
#   arrange(metric) %>% 
#   mutate(
#     value = ifelse(metric == "per",81,228),
#     taxon_emer = c(seq(2088,2099,1),seq(2088,2099,1))
#   )

plot_data <- plot_data %>% 
  # bind_rows(cont_eez) %>% 
  mutate(
    my_facet = ifelse(metric == "per","Percentage (%)","Cumulative sum (n stocks)"),
    label = ifelse(total == max(stock_d$com_sum),"Total number of emerging stocks","Total number of EEZs")
  )

```

```{r Time_lapse_plot, eval = F, echo = F}

### Three plots

count_eez_plot <- ggplot() +
  geom_area(data = subset(plot_data, metric == "com_sum" & level == "eez"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  )  +
  scale_y_continuous("Cumulative number of EEZs",
                     expand = c(0,0)
  ) +
  scale_x_continuous("Year of range shift",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100")
  ) +
  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
  scale_fill_viridis(discrete = T,
                     begin = 0.1, 
                     end = 0.1
  ) +
  facet_wrap(~ method);count_eez_plot

#### Count stock plot 
count_stock_plot <- ggplot() +
  geom_area(data = subset(plot_data, metric == "com_sum" & level == "stock"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  )  +
  scale_y_continuous("Cumulative number of stocks",
                     expand = c(0,0)
  ) +
  scale_x_continuous("",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100")
  ) +
  my_ggtheme_p(leg_pos = "", hjust = 0.5) +
  scale_fill_viridis(discrete = T,
                     begin = 0.7, 
                     end = 0.7
  ) +
  facet_wrap(~ method);count_stock_plot

# Percentage plot
percentage_plot <- ggplot() +
  geom_area(data = subset(plot_data, level == "eez" & my_facet == "Percentage (%)"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  ) + 
  geom_area(data = subset(plot_data, level == "stock" & my_facet == "Percentage (%)"),
            aes(
              x = as.numeric(taxon_emer),
              y = value,
              fill = level
            ),
            size=0.5,
            colour="black"
  ) + 
  scale_y_continuous("Cumulative proportion of the total (%)",
                     expand = c(0,0)
  ) +
  scale_x_continuous("",
                     breaks = seq(2000,2100,25),
                     limits = c(2005,2100),
                     label = c(seq(2000,2075,25),">2100")
  ) +
  my_ggtheme_p(leg_pos = "right", hjust = 0.5) +
  scale_fill_viridis(discrete = T,
                     begin = 0.1,
                     end = 0.7,
                     labels = c("EEZs","Stocks")
  )+
  facet_wrap(~ method);percentage_plot

# Legend
legend <- get_legend(percentage_plot)

percentage_plot <- percentage_plot + 
  theme(legend.position = "")

# Draw figure
plot_all <- ggdraw() +
  # Count stock plot
  draw_plot(count_stock_plot, x = 0.01, y = 0, width = 0.29, height = 0.90) +
  # Count EEZ plot
  draw_plot(count_eez_plot, x = 0.31, y = 0, width = 0.29, height = 0.90) +
  # Percentage plot
  draw_plot(percentage_plot, x = 0.61, y = 0 , width = 0.29, height = 0.90) +
  # legend
  draw_plot(legend, x = 0.84, y = 0 , width = 0.20, height = 1) +
  # Labels
  draw_plot_label(label = c("A", "B","C"), 
                  size = 20,
                  x = c(0.01, 0.285,0.59), 
                  y = c(1,1,1)
  )

# Save plot
save_plot(
  "./Figures/cumu_per_freq_sens.png",
  plot_all,
  base_height = 5,
  base_width = 12
)

```

#### Comparative map

```{r}

# Per EEZs
Eez_d_hist <- sens_data %>% 
  group_by(eezid,sau_fishing_entity,tresh,method) %>% 
  summarise(
    mean_yr = mean(emerging_yr,na.rm=T),
    sd_yr = sd(emerging_yr,na.rm=T),
    n_taxa = length(unique(taxon_key))
  ) %>% 
  gather("metric","value",mean_yr:n_taxa) %>% 
  mutate(
    reverse = abs(value-2100)  + 2000
  )

# Estimate the value (catch and USD) of emerging species
Eme_value_data <- sens_data %>% 
  group_by(sau_fishing_entity,fishing_entity_id,tresh,taxon_key,method) %>% 
  summarise(n()) %>% 
  left_join(Sau_data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  filter(!is.na(variable)) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,tresh,variable,method) %>% 
  summarise(
    sum_value = sum(value,na.rm=T),
    n_taxa = length(unique(taxon_key)) 
  )

# Total value of each fishing nation

Total_value <- Sau_data %>% 
  group_by(fishing_entity_id,variable) %>% 
  summarise(tot_value= sum(value,na.rm=T))

Value_map_data <- Eme_value_data %>% 
  left_join(Total_value,
            by = c("fishing_entity_id","variable")) %>% 
  mutate(participation = (sum_value/tot_value)*100+2000)

# Per fishing entity

fish_d_hist <- sens_data %>% 
  group_by(sau_fishing_entity,tresh,method) %>% 
  summarise(
    value = length(unique(taxon_key))
  ) %>% 
  filter(!is.na(sau_fishing_entity))

World_map_nemer <- World_map %>%
  full_join(Value_map_data,
            by = c("sau_fishing_entity","fishing_entity_id")
  ) %>% 
  filter(!is.na(participation))

# These are regions without transboundary species
grey_land <- World_map %>% 
  filter(!fishing_entity_id %in% Value_map_data$fishing_entity_id) %>% 
  filter(!sovereignt %in% c("Antarctica"))

# -------------------------- #
# Call plot
# -------------------------- #
tresh_list <- "tresh_two"

name_save <- paste("./Figures/emer_map_sens_",tresh_list,".png",sep = "")

# Value Map
World_map_part <- World_map_nemer %>% 
  filter(
    tresh ==  tresh_list,
    variable == "value"
  )

# ToE Map 
toe_map_data <- World_eez_sf %>% 
  left_join(Eez_d_hist,
            by = "eezid") %>% 
  filter(tresh == tresh_list,
         metric == "mean_yr")

# Map with countries without ToE

russia_churchil <- World_eez_sf %>% 
  filter(eez_name %in% c("Russia (Kara Sea)","Russia (Laptev to Chukchi Sea)"))

no_toe_eez <- World_eez_sf %>% 
  filter(eez_name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
         !eez_name %in% toe_map_data$eez_name) %>%  # Filter EEZs with ToE stocks
  rbind(russia_churchil)

no_toe_land <- World_map %>% 
  filter(name %in% Neighbours$eez_name, # Filter EEZs without Transboundary spp
         !name %in% World_map_part$name) # Filter EEZs with ToE stocks

map <- ggplot(toe_map_data) +
  geom_sf(
    aes(
      fill = reverse
    ),
    color =  "white",
    size = 0.1
  ) +
  geom_sf(data = World_map_part,
          aes(fill = participation),
          color =  "white",
          size = 0.1
  ) +
  geom_sf(data = grey_land, aes(), fill = "grey50", color =  "white",size = 0.1) +
  geom_sf(data = no_toe_eez, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  geom_sf(data = no_toe_land, aes(), fill = "lightblue", color =  "white",size = 0.1) +
  scale_fill_viridis("",
                     limits = c(2000,2100),
                     breaks = seq(2000,2100,25),
                     labels = c(">2100",seq(2075,2000,-25)) # not inluding the > because we
  ) +
  my_ggtheme_m(leg_pos = "bottom") +
  facet_wrap(~method,
             nrow = 2)

ggsave(name_save,
       width = 10,
       height = 6,
       units = "in"
)

```

# END PAPER

# Troubleshooting

### Weird connections from Australia

```{r fig_map_Juan, eval = F, echo =F}

##_----------------------- #
# Checking on China and Indonesia
# Probably due to dispute territories

Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
  clean_names()

# 
# Neighbours %>% 
#   filter(eez_name == "China") %>% 
#   pull(eez_neighbour) %>% 
#   unique()

##_----------------------- #

# Get the center poligon of each EEZ for map source/target
coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
  ungroup() %>% 
  select(eez_name,geometry) %>% 
  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
  # Manually fix issues
  mutate(
    longitude = ifelse(
      eez_name == "Fiji", 178.065033,
      ifelse(eez_name == "Tuvalu",177.6493,
             ifelse( eez_name == "New Zealand",174.8860,
                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
                     )
             )
      )
    )
  )

# Remove "c()"
coords$longitude <- gsub("\\(","",coords$longitude)
coords$longitude <- gsub("c","",coords$longitude)
coords$latitude <- gsub(")","",coords$latitude)

coords <- coords %>% 
  mutate_at(vars(longitude,latitude),as.numeric)

# Prepare data from results
Test_Data <- results %>% 
  # get coords for sources
  left_join(coords,
            by = "eez_name") %>%
  rename(source = eez_name,
         start_lon = longitude,
         start_lat = latitude,
         eez_name=eez_neighbour) %>% 
  # get coords for targets
  left_join(coords,
            by = "eez_name") %>% 
  rename(target = eez_name,
         end_lon = longitude,
         end_lat = latitude) %>% 
  ungroup() %>% 
  # Plot miscs
  mutate(
    per_mid = ifelse(is.na(per_mid),0,per_mid)
  ) %>% 
  drop_na() %>% 
  filter(
    # per_mid >= 20,
    variable == "mean_spp",
  )

# Convert to data to spatial
network_data_sf <- Test_Data %>%
  select(start_lon, start_lat, end_lon, end_lat) %>% 
  purrr::transpose() %>% 
  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
  purrr::map(st_linestring) %>%
  st_sfc(crs = 4326) %>%
  st_sf(geometry = .)%>%
  bind_cols(Test_Data) %>%
  select(everything(), geometry) %>% 
  # Set circles ('cus the world is not flat!)
  st_segmentize(units::set_units(100, km)) %>% 
  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  sf::st_sf(crs = 4326)

#  unpack the coordinates of each feature 
coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))

c(network_data_sf$line_id, network_data_sf$coords) %<-% (
  coord_data %>% 
    rename(subline_id = L1) %>%   
    nest(-L2)
)

# Create data path for line connections
paths <- network_data_sf %>% 
  sf::st_set_geometry(NULL) %>% 
  unnest() %>% 
  mutate_at(vars(X,Y), round,2)

# Plot it!

country <- c("Fiji","Solomon Isl.","Tuvalu")

ggplot() +
  geom_sf(
    # data = subset(world_map,name %in% country),
    data = world_map,
    size = 0.1, fill = "gray90", col = "gray90") +
  geom_sf(
    # data = subset(World_eez_sf, eez_name %in% country),
    data = World_eez_sf,
    size = 0.1, fill = "white", col = "gray70") +
  geom_point(data = Test_Data,
             # data = subset(Test_Data, source %in% country & target %in% source),
             aes(start_lon,
                 start_lat,
                 col = per_mid),
             size = 0.1,
             col = "black",
             # shape = 21,
             stroke = 1.5) +
  geom_path(data = paths,
            # data = subset(paths, source %in% country),
            aes(X, Y,
                group = interaction(line_id,subline_id),
                col = per_mid
            ),
            size = 0.3,
            arrow = arrow(length = unit(0.009, "npc")),
            show.legend = NA) +
  # geom_sf_text(data = subset(World_eez_sf, eez_name %in% country),
  #              aes(label =eez_name),
  #              size = 2) +
  # geom_text(data = subset(Test_Data, source %in% country),
  #              aes(start_lon,
  #                start_lat,
  #                label = paste(source,"to",target)
  #              ),
  #              size = 2) +
  # geom_text_repel(data = Test_Data,
  #   aes(start_lon,
# start_lat,
# label = paste(source,"to",target)
# )
# )+
ggtheme_map() +
  scale_colour_gradientn("Gain of Stock-Share Ratio (%)",
                         colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                         limits = c(0,25),
                         breaks = seq(0,25,5)
  )

```

### Arctic and Brasil missing species

```{r arctic, eval = F, echo =F}

# Old Code

###### Table.1 Multy linear model

Arctic <- c("Canada (Arctic)","USA (Alaska, Arctic)", "Russia (Laptev to Chukchi Sea)")

can_us_arctic <- Transboundary_spp %>% 
  filter(eez_name == "Canada (Arctic)",
         eez_neighbour == "USA (Alaska, Arctic)")

arctic_spp <- Transboundary_spp %>% 
  filter(eez_name %in% Arctic,
         eez_neighbour %in% Arctic)

arctic_spp <- Spp_depth %>% 
  filter(taxon_key %in% can_us_arctic$taxon_key)

# Transbounudary species that will see no change
# 600252 Mallotus villosus  (Capelin)
# 600315 Eleginus gracilis (Safforn cod)
# 601520 Clupea pallasii  (Pacific herring)

Eme_Res %>% 
  filter(taxon_key %in% can_us_arctic$taxon_key) %>% 
  # filter(sau_eez_name %in% arctic_spp$eez_name) %>% 
  View(.)

```

```{r Brasa, eval = F, echo =F}

brasil <- Transboundary_spp %>% 
  filter(eez_name == "Brazil")

brasil_spp <- Spp_depth %>% 
  filter(taxon_key %in% brasil$taxon_key)

# Transbounudary species that will see no change
# 601590  Brevoortia pectinata  
# 690022  Pleoticus muelleri 
# 690023  Illex argentinus  
# 690338  Penaeus paulensis

Eme_Res %>% 
  filter(taxon_key %in% brasil_spp$taxon_key) %>% 
  View(.)

```

```{r TransIndex, eval = F, echo = F}

TransIndex <- function(Spp,year, data_type = "Catch",ensemble_list,path,Neighbours,eez_centroid){

  # ----------------- #
  # Get Species Centrodis
  #  ---------------- #

  # Get model data from spps
  Spp_Dist <- bind_rows(
    lapply(
      ensemble_list,
      dbem_import,
      taxon_key = Spp,
      year = year,
      data_type = data_type,
      path = path
    )
  )

  #### Control for Spp id so I can fix problems

  if(file.exists("~/Desktop/spp_df.csv") == FALSE){
    spp_df <- tibble() %>% 
      write_csv(.,
                "~/Desktop/spp_df.csv")
  }

  suppressMessages(
    spp_df <- read_csv("~/Desktop/spp_df.csv")
  )
  spp_df <- tibble(Spp) %>% 
    bind_rows(spp_df) %>% 
    write_csv(.,
              "~/Desktop/spp_df.csv")

  #____________ Selecting only the transboundary nature of the species _________ #
  Trans_Spp <- Spp_Dist %>%
    left_join(index_code,
              by= "index") %>% 
    filter(!is.na(value)) %>% 
    semi_join(Transboundary_spp, # filter only transboundary cases
              by = c("taxon_key","eez_name")
    )

  # Step
  nr <- nrow(Trans_Spp)
  if(nr > 0){

    # Filter eez by those transboundary
    unique_eez <- unique(Trans_Spp$eez_name)

    Neighbours_combo <- Neighbours %>% 
      filter(eez_name %in% unique_eez,
             eez_neighbour %in% unique_eez)

    # Get the centroid of each country
    EEZ_centroid <- eez_centroid %>% 
      filter(eez_name %in% Trans_Spp$eez_name) %>% 
      mutate(taxon_key = Spp)

    Ids <- unique(Neighbours_combo$neighbour_id)

    corefx <- function(Ids){

      EEZ_ids <- Neighbours_combo %>% 
        filter(neighbour_id == Ids)

      Sub_Trans_Spp <- Trans_Spp %>% 
        filter(eez_name %in% EEZ_ids$eez_name)

      Sub_EEZ_centroid <- EEZ_centroid %>% 
        filter(eez_name %in% EEZ_ids$eez_name)

      # Centroid is defined as the tile with the higest species value
      Centroids <- Sub_Trans_Spp %>% 
        left_join(EEZ_ids, by = "eez_name") %>% 
        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
        top_n(1,value) %>% 
        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
        select(eez_name =eez_name.y, everything() ,-eez_name.x)

      # Estimate distances between spp centroid and EEZ centroid
      Distance_Data <- Centroids %>%
        rename(
          catch_lon = lon,
          catch_lat = lat,
          eez_lon = longitude,
          eez_lat= latitude
        ) %>% 
        # Estimate the distance between centriods
        mutate(
          distance = geodist(eez_lon, eez_lat, catch_lon, catch_lat, units="km")
        ) %>% 
        select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)

      # Historical mean and sd of distance
      historical <- Distance_Data %>%
        filter(year <= 2000) %>%
        ungroup() %>%
        group_by(taxon_key,ensemble,eez_name,neighbour_id) %>%
        summarise_at(vars("distance"),
                     c(hd_temp_mean=mean,
                       hd_temp_sd=sd)
        )

      # Future results
      future <- Distance_Data %>% 
        filter(year > 2000)

      # Transboundary Index
      Trans_index <- Distance_Data %>% 
        left_join(historical,
                  by = c("taxon_key","eez_name","ensemble","neighbour_id")
        ) %>%
        # Re arrenging table for mutate
        gather("variable","value",distance:hd_temp_sd) %>%
        ungroup()

      Partial <- Trans_index %>% 
        mutate(
          names = ifelse(eez_name ==Sub_EEZ_centroid$eez_name[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_"))
        ) %>%
        select(-eez_name,-variable) %>%
        spread(names,value) %>% 
        # mutate index
        mutate(
          trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
        ) %>% 
        group_by(taxon_key,ensemble,neighbour_id) %>% 
        mutate(RMean = rollmean(x = trans_index, 
                                10, 
                                align = "right", 
                                fill = trans_index)
        ) %>% 
        filter(year > 1960) %>% 
        select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)

      # Average of 10 years mean of ensemble members results
      mean_sd_index <- Partial %>% 
        group_by(taxon_key,year,neighbour_id) %>% 
        summarise_at(vars("RMean"),
                     c(trend=mean,
                       ensemble_sd=sd)
        )

      # Estimate noice from ensemble members
      historic_sd <- mean_sd_index %>% 
        filter(year < 2000) %>% 
        group_by(taxon_key) %>% 
        summarise_at(vars("trend"),
                     c(hist_mean=mean,
                       noise=sd)
        )

      # Plot the result, bruh
      # ggplot() +
      #   geom_line(data = mean_sd_index,
      #             aes(
      #               x = year,
      #               y = trend
      #             )
      #   ) +
      #   geom_ribbon(data = mean_sd_index,
      #               aes(x=year,
      #                   ymin=historic_sd$hist_mean-historic_sd$noise,
      #                   ymax=historic_sd$hist_mean+historic_sd$noise),
      #               alpha = 0.5,
      #               fill = "grey50"
      #   ) +
      #   geom_line(data = mean_sd_index,
      #             aes(
      #               x = year,
      #               y = historic_sd$hist_mean
      #             ),
      #             colour = "grey50"
      #   ) +
      #   geom_point(data = subset(mean_sd_index, trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise)),
      #              aes(
      #                x = year,
      #                y = trend,
      #                colour = "red",
      #              ),
      #              shape = 1,
      #              size = 3
      #   ) +
      #   geom_vline(xintercept = 2000,colour="blue",alpha = 0.5) +
      #   theme_classic()

      # #     
      Final_Result <- mean_sd_index %>%
        filter(year > 2000,
               trend > (historic_sd$hist_mean+historic_sd$noise) |
                 trend < (historic_sd$hist_mean-historic_sd$noise)
        ) %>%
        group_by(taxon_key,neighbour_id) %>%
        summarise(emerging_yr= min(year)) %>%
        left_join(EEZ_ids) %>%
        ungroup() %>%
        select(taxon_key,eez_name,eez_neighbour,emerging_yr)

      # if(i == 1){
      #   f_df <- Final_Result
      # }else{
      #   f_df <- bind_rows(f_df,Final_Result)
      # }
      return(Final_Result)
    }

    x <- bind_rows(
      mclapply(Ids,corefx)
    )
    # return(f_df)

  }else{

    print("no share data")

    x <- tibble(
      taxon_key = Spp,
      eez_name = NA,
      emerging_yr = NA
    )

  }

  return(x)

}

#  #One spp works 
# suppressMessages(
# TransIndex(Spp = 600006,
#              year,
#              data_type = "Catch",
#              ensemble_list,
#              path,
#              Neighbours = Neighbours,
#              eez_centroid)
# Multiple spp

# spp_list <- c(600004,
#               600005,
#               600245,
#               600006
#               )
# 
# suppressMessages(
#   suppressWarnings(
#     Test <- bind_rows(
#       lapply(spp_list,
#              TransIndex,
#              year,
#              data_type = "Catch",
#              ensemble_list,
#              path,
#              Neighbours,
#              eez_centroid
#       )
#     )
#   )
# )

```

###### With relation to neighbours

```{r time_emergence_map,  eval = F, echo = T}

##_----------------------- #
# Checking on China and Indonesia
# Probably due to dispute territories

Neighbours <- my_path("D", extra_path = "Spatial/", name="Neighbours_eez_id.csv", read = TRUE) %>% 
  clean_names()

# 
# Neighbours %>% 
#   filter(eez_name == "China") %>% 
#   pull(eez_neighbour) %>% 
#   unique()

##_----------------------- #

# Get the center poligon of each EEZ for map source/target
coords <- as.data.frame(st_centroid(World_eez_sf)) %>% 
  ungroup() %>% 
  select(eez_name,geometry) %>% 
  separate(col = geometry, into = c("longitude", "latitude"), sep = "\\,") %>% 
  # Manually fix issues
  mutate(
    longitude = ifelse(
      eez_name == "Fiji", 178.065033,
      ifelse(eez_name == "Tuvalu",177.6493,
             ifelse( eez_name == "New Zealand",174.8860,
                     ifelse( eez_name == "USA (Alaska, Subarctic)",-145.8860,longitude
                     )
             )
      )
    )
  )

# Remove "c()"
coords$longitude <- gsub("\\(","",coords$longitude)
coords$longitude <- gsub("c","",coords$longitude)
coords$latitude <- gsub(")","",coords$latitude)

coords <- coords %>% 
  mutate_at(vars(longitude,latitude),as.numeric)

# Prepare data from results
Test_Data <- Eme_Res %>% 
  filter(
    n_ens == 10#,
    # emerging_yr >= 2020
  ) %>% 
  group_by(eez_name,eez_neighbour,emerging_yr) %>% 
  summarise(
    n_tax = length(unique(taxon_key))
  ) %>% 
  filter(
    emerging_yr <= 2050
  ) %>%
  group_by(eez_name,eez_neighbour) %>% 
  summarise(
    n_tax_emer = sum(n_tax)
  ) %>% 
  # get coords for sources
  left_join(coords,
            by = "eez_name") %>%
  rename(source = eez_name,
         start_lon = longitude,
         start_lat = latitude,
         eez_name=eez_neighbour) %>% 
  # get coords for targets
  left_join(coords,
            by = "eez_name") %>% 
  rename(target = eez_name,
         end_lon = longitude,
         end_lat = latitude) %>% 
  ungroup() %>% 
  # Plot miscs
  # mutate(
  #   per_mid = ifelse(is.na(per_mid),0,per_mid)
  # ) %>% 
  drop_na() #%>% 
# filter(
# per_mid >= 20,
# variable == "mean_spp",
# )

# Convert to data to spatial
network_data_sf <- Test_Data %>%
  select(start_lon, start_lat, end_lon, end_lat) %>% 
  purrr::transpose() %>% 
  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
  purrr::map(st_linestring) %>%
  st_sfc(crs = 4326) %>%
  st_sf(geometry = .)%>%
  bind_cols(Test_Data) %>%
  select(everything(), geometry) %>% 
  # Set circles ('cus the world is not flat!)
  st_segmentize(units::set_units(100, km)) %>% 
  mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  sf::st_sf(crs = 4326)

#  unpack the coordinates of each feature 
coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))

c(network_data_sf$line_id, network_data_sf$coords) %<-% (
  coord_data %>% 
    rename(subline_id = L1) %>%   
    nest(-L2)
)

# Create data path for line connections
paths <- network_data_sf %>% 
  sf::st_set_geometry(NULL) %>% 
  unnest() %>% 
  mutate_at(vars(X,Y), round,2)

# Plot it!

country <- c("Fiji","Solomon Isl.","Tuvalu")

ggplot() +
  geom_sf(
    # data = subset(world_map,name %in% country),
    data = World_map,
    size = 0.1, fill = "gray90", col = "gray90") +
  geom_sf(
    # data = subset(World_eez_sf, eez_name %in% country),
    data = World_eez_sf,
    size = 0.1, fill = "white", col = "gray70") +
  geom_point(data = Test_Data,
             # data = subset(Test_Data, source %in% country & target %in% source),
             aes(start_lon,
                 start_lat,
                 col = n_tax_emer
             ),
             size = 0.1,
             col = "grey50",
             # shape = 21,
             stroke = 1.5) +
  geom_path(data = paths,
            # data = subset(paths, source %in% country),
            aes(X, Y,
                group = interaction(line_id,subline_id),
                col = n_tax_emer
            ),
            size = 0.3,
            # arrow = arrow(length = unit(0.009, "npc")),
            show.legend = NA) +
  ggtheme_map() +
  scale_colour_viridis(
    name="Number of emerging stocks\nby 2050",
    limits = c(0,40),
    breaks = seq(0,40,10),
    # option="magma",
    direction = 1
  ) +
  ggsave("./Figures/figure_emerging_2050.png",
         width = 10,
         height = 6,
         units = "in"
  )

```

### Non selected Results

```{r, olda_acumulative, eval = F}

# filter(emerging_yr >= 2020) %>% 
ggplot() +
  geom_area(data = subset(plot_data, level == "stock"),
            aes(
              x = as.numeric(taxon_emer),
              y = com_sum,
              fill = level
            ),
            alpha= 0.8,
            size=0.5,
            colour="black"
  )  +
  geom_area(data = subset(plot_data, level == "eez"),
            aes(
              x = taxon_emer,
              y = com_sum,
              fill = level
            ),
            alpha= 0.8,
            size=0.5,
            colour="black"
  ) +
  facet_wrap(~tresh) + 
  # Comment out for option A (Change plot name!!)
  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
  scale_y_continuous("Cumulatuve Curve",
                     breaks = seq(0,600,50),
                     limits = c(0,600)
  ) +
  scale_x_continuous("Year of emergence",
                     breaks = seq(2005,2100,10),
                     limits = c(2005,2100)
  ) +
  my_ggtheme_p() +
  scale_fill_viridis(discrete = T) +
  ggsave("./Figures/figure_accumulative_curve_ab.png",
         width = 6,
         height = 6,
         units = "in" 
  )

### Percentage accumulation

ggplot() +
  geom_area(data = subset(plot_data, level == "stock"),
            aes(
              x = as.numeric(taxon_emer),
              y = per,
              fill = level
            ),
            alpha= 0.8,
            size=0.5,
            colour="black"
  )  +
  geom_area(data = subset(plot_data, level == "eez"),
            aes(
              x = taxon_emer,
              y = per,
              fill = level
            ),
            alpha= 0.8,
            size=0.5,
            colour="black"
  ) +
  scale_y_continuous("Cumulatuve Curve",
                     breaks = seq(0,100,25),
                     limits = c(0,100)
  ) +
  scale_x_continuous("Year of emergence",
                     breaks = seq(2005,2100,15),
                     limits = c(2005,2100)
  ) +
  my_ggtheme_p() +
  scale_fill_viridis(discrete = T) +
  # facet_wrap(~tresh) + 
  ggsave("./Figures/Fig4.png",
         width = 8,
         height = 8,
         units = "in"
  )

```

```{r histo,  eval = F, echo = T}

# First time a Species becomes emergent
Eme_Res %>% 
  filter(
    n_ens == 10#,
    # tresh == "tresh_one"
  ) %>% 
  group_by(tresh,taxon_key,emerging_yr) %>% 
  summarise(
    n_trans = length(unique(sf_eez_name))
  ) %>% 
  group_by(tresh,taxon_key) %>% 
  summarise(
    ntrans = min(emerging_yr)
  ) %>% 
  group_by(tresh,ntrans) %>% 
  summarise(
    n_taxa = length(unique(taxon_key))
  ) %>% 
  arrange(-n_taxa) %>% 
  # View()
  # group_by(tresh) %>% 
  # summarise(sum(n_taxa))
  ggplot() +
  geom_line(
    aes(
      x = ntrans,
      y = n_taxa,
      color = tresh
    )
  )

# EEZ, first time an EEZ has a emergent stock
Eme_Res %>% 
  filter(
    n_ens == 10#,
    # tresh == "tresh_one"
  ) %>% 
  group_by(tresh,sf_eez_name,emerging_yr) %>% 
  summarise(
    n_trans = length(unique(taxon_key))
  ) %>% 
  group_by(tresh,sf_eez_name) %>% 
  summarise(
    ntrans = mean(emerging_yr)
  ) %>% 
  group_by(tresh,ntrans) %>% 
  summarise(
    n_taxa = length(unique(sf_eez_name))
  ) %>% 
  arrange(-n_taxa) %>% 
  View()
# group_by(tresh) %>%
# summarise(sum(n_taxa))
# ggplot() +
# geom_line(
#   aes(
#     x = ntrans,
#     y = n_taxa,
#     color = tresh
#   )
# )

# Numer of EEZs where stocks will emerge
stock_d_hist <- Eme_Res %>% 
  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
  group_by(emerging_yr,taxon_key,tresh) %>%
  summarise(n_tax = length(unique(sf_eez_name))) %>%
  # summarise(n()) %>% 
  mutate(
    level = "EEZs per Spp"
  )  %>% 
  ungroup() %>% 
  select(-taxon_key)

# Per EEZs
eez_d_hist <- Eme_Res %>% 
  distinct(taxon_key,neighbour_id,tresh, .keep_all = TRUE) %>%
  group_by(emerging_yr,sf_eez_name,tresh,neighbour_id) %>%  
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(n()) %>% 
  mutate(
    level = "Spp per EEZ"
  ) %>% 
  ungroup() %>% 
  select(-sf_eez_name)

stock_d_hist %>% 
  bind_rows(eez_d_hist) %>% 
  filter(tresh != "NA") %>% 
  ggplot() +
  geom_density(
    aes(
      x = emerging_yr,
      fill = level
    ),
    stat = "count",
    alpha= 0.8,
    size=0.5,
  ) +
  scale_y_continuous("Frequency",
                     # breaks = seq(0,300,50),
                     # limits = c(0,300)
  ) +
  scale_x_continuous("Year of emergence",
                     # breaks = seq(2006,2100,25),
                     # limits = c(2000,2100)
  ) +
  my_ggtheme_p() +
  scale_fill_viridis(discrete = T) +
  facet_wrap(~tresh) +
  ggsave("./Figures/hist_freq.png",
         width = 6,
         height = 6,
         units = "in"
  )

```

##### Temperature gradient

##### Fun `GetEnvData`

```{r get_env_data, eval = F, echo = T}

getEnvVar=function(yr,ModelRCP,RCP,Variable){

  if(ModelRCP == "IPSL"){
    ModelRCP = "IPSL-CM5A-MR"
    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
  }

  if(ModelRCP == "MPI"){
    ModelRCP = "MPI-ESM-MR"
    Filepath <- paste("/Volumes/DATA/DATA/Environmental data/Model Climate data/CMIP5/txt CMIP5/Reformatted",ModelRCP,RCP,"",sep="/")
  }

  if(ModelRCP == "GFDL"){

    if(RCP == "rcp26"){
      ModelRCP <- "GFDL26"
    }else{
      ModelRCP <- "GFDL85"
    }
    Filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")
  }

  if(Variable == "Ice"){

    Bottom <- paste("IceExt_",yr,".txt",sep="")

  }

  if(Variable == "Temperature"){

    Bottom = paste("SST_",yr,".txt",sep="")

  }

  if(Variable == "Salinity"){
    # Bottom
    Bottom = paste("Salinity_btm_",yr,".txt",sep="")

  }

  Read_me <- paste(Filepath,Bottom,sep="")[1]
  Bottom_Data <- read.csv(Read_me,header=F)

  # Retrun

  getTemp <- Bottom_Data 
}

Years <-seq(1951,2100,1)

for(y in 1:length(Years)){
  Data <- getEnvVar(yr = Years[y],
                    Model = "GFDL",
                    Variable = "Temperature",
                    RCP = "rcp85"
  )

  if(y == 1){

    F_BData <- Data[1]
    colnames(F_BData) <- Years[1]

  }else{

    F_BData <- cbind(F_BData,Data[1])
    colnames(F_BData)[y] <- Years[y]
  }
}

F_BData <- F_BData %>% 
  rowid_to_column("index") %>% 
  gather("year","temp",-index) %>% 
  mutate(temp = ifelse(temp == -9999,NA,temp)) %>% 
  group_by(year) %>% 
  summarise(
    mean_temp = mean(temp,na.rm=T)
  ) %>% 
  write_csv(.,
            "SST_data.csv")

```

```{r temp_plot, eval = F, echo = T}

### NEED TO RUN TIME LAPSE CHUNK FIRST TO GET ACCUMULATIVE DATA

# Get sst_data
sst_data <- my_path("D",T,"Climate","SST_data.csv",read = T)

sst_hist <- sst_data %>% 
  filter(year < 2000) %>% 
  group_by() %>% 
  summarise(hist_sst =mean(mean_temp)) %>% 
  pull()

sst_change <- sst_data %>%
  filter(year > 2020) %>% 
  mutate(chng = round(mean_temp-sst_hist,2))

# Plot Data
stock_d %>% 
  bind_rows(eez_d) %>% 
  filter(taxon_emer >= 2020) %>% 
  rename(year = taxon_emer) %>% 
  left_join(sst_change) %>% 
  ggplot() +
  geom_line(
    aes(
      x = chng,
      y = com_sum,
      colour = level
    )
  ) +
  geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
  geom_text(aes(label="Total number of EEZs", x = 0.5, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
  scale_y_continuous("Cumulatuve Curve",
                     breaks = seq(0,450,50),
                     limits = c(0,500)
  ) +
  scale_x_continuous("Change in SST (C)",
                     breaks = seq(0,2,0.2),
                     limits = c(0.3,2.3)
  ) +
  my_ggtheme_p() +
  scale_colour_viridis(discrete = T) +
  ggsave("./Figures/figure_temp.png",
         width = 6,
         height = 6,
         units = "in"
  )

```

```{r Scatter_plot,  eval = F, echo = T}

data_stock <- Eme_Res %>% 
  filter(
    n_ens == 10,
    emerging_yr >= 2020
  ) %>% 
  group_by(emerging_yr,eez_name) %>% 
  summarise(
    tax_n = length(unique(taxon_key))
  ) %>% 
  group_by(emerging_yr) %>% 
  summarise(emerging_tax = sum(tax_n))

data_eez <- Eme_Res %>% 
  filter(
    n_ens == 10,
    emerging_yr >= 2020
  ) %>% 
  group_by(emerging_yr,taxon_key) %>% 
  summarise(
    eez_n = length(unique(eez_name))
  ) %>% 
  group_by(emerging_yr) %>% 
  summarise(emerging_eez = sum(eez_n))

data_stock %>% 
  left_join(data_eez,
            by = "emerging_yr") %>% 
  rename(year = emerging_yr) %>% 
  left_join(sst_data) %>% 
  ggplot() +
  geom_point(
    aes(
      x = mean_temp,
      y = emerging_eez,
      fill = year,
      colour = year
    )
  )

```

# Old code

## Fig. 4. Bubble plot

```{r, buble_plot, eval = F, echo = F}

# Needs previous chunk

Buble_data <- Eez_d_hist %>% 
  filter(metric == "mean_yr") %>% 
  group_by(sau_fishing_entity,tresh) %>% 
  summarise(mean_yr = mean(value)) %>% 
  left_join(Eme_value_data,
            by = c("sau_fishing_entity","tresh")) %>% 
  filter(#tresh == selected_tresh, # tresh hold selection
    variable == "value") %>% 
  mutate(
    weight_toe = sum_value/mean_yr,
    sum_value_bi = sum_value/1000000000,
    log_value = log10(sum_value_bi)
  ) %>%   # Billion usd
  left_join(Matching_names,
            by = c("sau_fishing_entity","fishing_entity_id")) %>% 
  distinct(sau_fishing_entity,tresh,.keep_all = TRUE) %>%
  filter(!is.na(un_geo_region)) %>% 
  ungroup()# %>%  # no more asterisk
# mutate(
#   sau_fishing_entity = ifelse(sum_value_bi >= 1 & !sau_fishing_entity %in% c("Spain"), paste(sau_fishing_entity,"*",sep = ""),sau_fishing_entity))

### Function for estimating percentile

quibble <- function(x, q = p) {
  tibble(x = quantile(x, q), q = q)
}

Percentile_value <- Buble_data %>% 
  group_by(tresh) %>%  # for whole distribution
  do(quibble(.$log_value, 0.75))

## Top fishing nations close to SDGs
Buble_data %>% 
  filter(log_value > Percentile_value$x & mean_yr < 2036) %>% 
  View()
group_by(tresh) %>% 
  summarise(sum(sum_value_bi),
            sum(n_taxa))

## All fishing nations close to SDGs
Buble_data %>% 
  filter(mean_yr < 2036) %>% 
  # nrow() # 76  countries
  View()
group_by(tresh) %>% 
  summarise(sum(sum_value_bi),
            sum(n_taxa))

76/length(unique(Buble_data$sau_fishing_entity))
134/length(unique(Buble_data$sau_fishing_entity))

# Labels Data

Labels_data <- tibble(
  x = c(2029,
        # 2050,2070,
        2007),
  y = c(-5.2,
        # -5.8,-5.8,
        Percentile_value$x+.1),
  label = c("UN 2030 Agenda", 
            # "Mid shift","Late shift",
            "Fisheries > 75th percentile")
)

# Plot the bubbles!
ggplot(Buble_data) +
  geom_point(
    aes(
      x = mean_yr,
      y = log_value,
      size = n_taxa,
      colour = un_geo_region,
      shape = un_geo_region
    ),
    alpha = 0.7
  ) +
  # Right side labels
  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034),
                   aes(
                     x = mean_yr,
                     y = log_value,
                     label = sau_fishing_entity
                   ),
                   arrow = arrow(length = unit(0.009, "npc"),
                                 type = "closed",
                                 ends = "first"),
                   force = 1,
                   box.padding = 0.2,
                   fill = "NA",
                   nudge_x = 2066 - subset(Buble_data,log_value > Percentile_value$x & mean_yr < 2036 & mean_yr > 2034)$mean_yr, # set distance from point
                   nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
                   segment.size  = 0.1,
                   segment.color = "grey50",
                   direction     = "y",
                   hjust         = 1
  ) +
  # LEFT side labels
  geom_label_repel(data = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034),
                   aes(
                     x = mean_yr,
                     y = log_value,
                     label = sau_fishing_entity
                   ),
                   arrow = arrow(length = unit(0.009, "npc"),
                                 type = "closed",
                                 ends = "first"),
                   force = 1,
                   box.padding = 0.2,
                   fill = "NA",
                   nudge_x = 2025 - subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$mean_yr, # set horizontal position
                   nudge_y = subset(Buble_data, log_value > Percentile_value$x & mean_yr < 2034)$log_value  +1.1 , # set vertical position
                   segment.size  = 0.1,
                   segment.color = "grey50",
                   direction     = "y",
                   hjust         = 1
  ) +
  # Divisory lines and labels
  geom_vline(xintercept = 2030, color = "grey50", size = 0.5, linetype = "dashed") +
  geom_text(data = subset(Labels_data, x == 2029),
            aes(
              x = x,
              y = y,
              label = label
            ),
            size = 4,
            color = "grey30",
            angle = 90
  ) +
  geom_hline(data = Percentile_value,
             aes(yintercept = x),
             color = "grey50"
  )  +
  geom_text(data = subset(Labels_data, x == 2007),
            aes(
              x = x,
              y = y,
              label = label
            ),
            size = 4,
            color = "grey30"
  ) + 
  scale_shape("Region") +
  scale_color_viridis(name="Region",
                      discrete = T) +
  scale_x_continuous("Average year of range shift",
                     limits = c(2000,2080),
                     breaks = seq(2000,2080,10)
  ) +
  scale_y_continuous("Fishing revenue (log10)",
                     limits = c(-6,2),
                     breaks = seq(-6,2,2),
                     expand = c(0,0)
  ) +
  scale_size("Number of\nshifting stocks") +
  my_ggtheme_p(leg_pos = "right") +
  ggsave("./Figures/bubble_plot.png",
         width = 12,
         height = 8,
         units = "in"
  )

```

```{r participation_text, eval = T, echo=F}

Value_map_data %>% 
  group_by(tresh,variable) %>% 
  summarise(m = mean(participation)-2000,
            sd = sd(participation-2000)
  )

```

## Fig. 4N Top fishers on bars

The bar option by UN sub region of Figure 4N of top fishers

```{r}

# Create data
Species_Data <- top_toe %>% 
  filter(tresh == selected_tresh,
         variable == "value") %>% 
  mutate(mean = mean-2000,
         group = geo_region)

# Set a number of 'empty bar' to add at the end of each group
Char <- as.factor(Species_Data$group)
empty_bar <- 5
to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
colnames(to_add) <- colnames(Species_Data)
to_add$group <- rep(levels(Char), each=empty_bar)
Species_Data <- rbind(Species_Data, to_add)
Species_Data <- Species_Data %>% arrange(group,mean)
Species_Data$id <- seq(1, nrow(Species_Data))

# Species_Data$id <- seq(1, nrow(Species_Data))

# Get the name and the y position of each label
label_data <- Species_Data 
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# prepare a data frame for base lines
base_data <- Species_Data %>% 
  group_by(group) %>% 
  summarize(start=min(id), 
            end=max(id) - empty_bar, 
            median = median(mean, na.rm = T)
  ) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
# grid_data <- grid_data[-1,]
# grid_data[1,2] <- 5 # Manually include bathydemerssal
# grid_data[1,3] <- 1

un_regions <- grid_data %>% 
  select(group)

ggplot(Species_Data, 
       aes(x=as.factor(id), 
           y=mean,
           fill=mean)
) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(aes(x=as.factor(id),
               y=mean,
               fill=mean),
           stat="identity",
           alpha=0.5
  ) +
  geom_segment(data=grid_data,
               aes(x = end, y = 30, xend = start, yend = 30),
               colour = "grey90", alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_bar(aes(x=as.factor(id),
               y=mean,
               fill=mean
  ),
  stat="identity",
  alpha=0.5) +
  annotate("text", 
           x = 0.5,
           y = 37,
           label = "UN Agenda",
           color= "grey60", 
           size=3, 
           angle=-5
  ) +
  # Country names
  geom_text(data=label_data,
            aes(x=id, 
                y= 100,
                label=sau_fishing_entity,
                hjust=hjust),
            color="black",
            fontface="italic",
            alpha = 1,
            size = 4 ,
            angle = label_data$angle,
            inherit.aes = FALSE) +
  # Continents labels
  # geom_text(data=base_data,
  #           aes(x = title,
  #               y = 85, 
  #               label=group), 
  #           # hjust=c(1,1,1,0,0,0),
  #           colour = "black",
  #           # angle = myAng,
  #           size=5,
  #           fontface="bold",
  #           inherit.aes = FALSE) +
ylim(-100,100) + # sets the circle the first value the central the last value how wide
  theme_minimal() +
  theme(
    legend.position = c(0.5,0.5),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14, face = "bold"),
    legend.title.align = 0.5,
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(colour = "white"),
    legend.direction="horizontal",
    legend.key.width = unit(3, 
                            "line")
  ) +
  coord_polar(start = 0) +
  scale_fill_viridis("Average year of range shift\nfor top 5 most valuable stocks",
                     limits = c(0,100),
                     breaks = seq(0,100,20),
                     labels = seq(2000,2100,20),
                     direction = -1
  ) +
  guides(
    fill=guide_colorbar(title.position="top", 
                        title.hjust =0.5)
  )

ggsave(
  plot = last_plot(),
  width = 13,
  height = 13,
  units = "in",
  filename = "./Figures/top_stocks_circular_trend.png"
)

```

## Figure 5 code from Juan

```{r fig_map_Juan, eval = F, echo =F}
# Figure made thanks to Juanito!
# http://jsmayorga.com/post/mapping-the-global-network-of-transnational-fisheries/

### Median number of stocks changing in each time step
Prop_results %>% 
  filter(variable == "n_per") %>% 
  group_by() %>% 
  summarise_at(vars(early,mid),funs(
    median(.,na.rm=T),
    sd(.,na.rm=T),
    length(unique(eez_name)),
    length(unique(eez_name))/280*100
  )
  )

# Prepare data from results
Data <- Prop_results %>% 
  # get coords for sources
  left_join(Eez_centroid,
            # left_join(Eez_centroid_eck,
            by = "eez_name") %>%
  rename(source = eez_name,
         start_lon = longitude,
         start_lat = latitude,
         eez_name=eez_neighbour) %>% 
  # get coords for targets
  # left_join(Eez_centroid_eck,
  left_join(Eez_centroid,
            by = "eez_name") %>%
  rename(target = eez_name,
         end_lon = longitude,
         end_lat = latitude) %>% 
  ungroup() %>% 
  # Plot miscs
  mutate(
    per_early = ifelse(is.na(per_early),0,per_early),
    per_mid = ifelse(is.na(per_mid),0,per_mid)
  ) %>% 
  drop_na() %>%
  # gather("variable","value",per_early:per_early_plot) %>% 
  filter(
    variable %in% c("mean_eez","n_spp")#,
    # source %in% c("Brazil","Argentina","Chile","Peru","Uruguay")
  )

# Convert to data to spatial
network_data_sf <- Data %>%
  select(start_lon, start_lat, end_lon, end_lat) %>% 
  purrr::transpose() %>% 
  purrr::map(~ matrix(flatten_dbl(.), nrow = 2, byrow = TRUE)) %>% 
  purrr::map(st_linestring) %>%
  # st_sfc(crs = 4326) %>%
  st_sfc(crs = "+proj=eck4") %>% # for tubular plot
  st_sf(geometry = .) %>%
  bind_cols(Data) %>%
  select(everything(), geometry) %>% 
  # Set circles ('cus the world is not flat!)
  # st_segmentize(units::set_units(100, km)) %>%
  # mutate(geometry = (geometry + c(18000000,90000000)) %% c(18000000) - c(18000000,90000000)) %>%
  # mutate(geometry = (geometry + c(180,90)) %% c(360) - c(180,90)) %>%
  # st_wrap_dateline(options = "WRAPDATELINE=YES", quiet = TRUE); network_data_sf # for tubular plot
  st_wrap_dateline(options = c("WRAPDATELINE=YES",  "DATELINEOFFSET=60"), quiet = TRUE)

network_data_sf

#  unpack the coordinates of each feature 
coord_data <- as.data.frame( st_coordinates(st_cast(network_data_sf,"MULTILINESTRING")$geometry))

c(network_data_sf$line_id, network_data_sf$coords) %<-% (
  coord_data %>% 
    rename(subline_id = L1) %>%   
    nest(-L2)
)

# Create data path for line connections
paths <- network_data_sf %>% 
  sf::st_set_geometry(NULL) %>% 
  unnest(cols = c(coords)) %>%
  mutate_at(vars(X,Y), round,2) %>% 
  gather("time_step","per_chng",per_early:per_mid) %>% 
  mutate(per_plot = ifelse(per_chng > 50,50,per_chng),
         direction = ifelse(per_chng > 0,"positive","negative")
  )

# Plot it!

# Proportion map data 
Map_data <- Data %>% 
  gather("time_step","per_chng",per_early:per_mid) %>% 
  mutate(per_plot = ifelse(per_chng > 50,50,per_chng))

# Average number of taxa to change per eez

# Total number of transboundary stocks per EEZ
n_trans_eez <- Transboundary_spp %>% 
  group_by(eez_name) %>% 
  summarise(
    # n_trans = unique(length(taxon_key))
    n_trans = n()
  )

# Estimate the percentage changing
n_taxa_data <- Prop_results %>% 
  filter(variable == "n_spp") %>%
  select(eez_name,early,mid) %>% 
  left_join(n_trans_eez) %>% 
  # mutate(per = (early/n_trans)*100) %>% 
  mutate_at(vars(early,mid),
            .funs = ~(./n_trans)*100) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(Matching_names) %>% 
  select(sau_fishing_entity,fishing_entity_id,early,mid) %>% 
  gather("time_step","avrg_spp_per",early:mid) %>% 
  group_by(sau_fishing_entity,fishing_entity_id,time_step) %>%
  summarise(avrg_spp_per = mean(avrg_spp_per,na.rm=T)) %>% 
  mutate(time_step = ifelse(time_step == "early","per_early","per_mid")) %>%
  mutate(avrg_spp_per_100 = avrg_spp_per,
         avrg_spp_per = ifelse(avrg_spp_per > 50,50,avrg_spp_per))

# Paper numbers for Rev. 1

n_taxa_data %>% 
  group_by(time_step) %>% 
  summarise_at(vars(avrg_spp_per),
               c(mean,sd,min,max),
               na.rm=T)

# Included in map
n_taxa_map <- World_map %>% 
  full_join(n_taxa_data,
            by = c("sau_fishing_entity","fishing_entity_id")
  )

# Grey land
grey_land <- World_map %>% 
  filter(is.na(fishing_entity_id)) %>% 
  filter(!sovereignt %in% c("Antarctica"))

# Rectangles dataset
# coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
# Rectangles <- tibble(
#     x1=c(-100,-14), #LatAm, Med, NE
#     x2=c(-55,23),
#     y1=c(9,50), 
#     y2=c(30,67)
#   )

time_sets <- c("per_early","per_mid")
i = 1
# for(i in 1:2){
gc()# clean ram

time <- time_sets[i]

## ------------------- ##
# Global Plot
## ------------------- ##

Prop_map <- ggplot() +
  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
          aes(fill = avrg_spp_per),
          size = 0.1,
          col = "white") +
  geom_sf(data = World_eez_sf, size = 0.1, fill = "white", col = "gray30") +
  geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
  geom_point(data = subset(Map_data, variable =="mean_eez" & time_step == time & per_plot < 0), 
             aes(start_lon,
                 start_lat
             ),
             size = 0.05,
             col = "gray80",
             shape = 20,
             stroke = 1.5) +
  geom_path(data = subset(paths, variable =="mean_eez" & time_step == time & per_plot > 0), # filter data 
            aes(X,
                Y,
                group = interaction(line_id,subline_id),
                col = per_plot
            ),
            size = 1,
            lineend = "round",
            arrow = arrow(length = unit(0.008, "npc"),
                          ends = "first"),
            show.legend = NA) +
  my_ggtheme_m(leg_pos = "bottom") +
  scale_colour_viridis(
    name= "Changes in catch proportion \nand changing stocks (%)",
    limits = c(0,max(Map_data$per_plot)),
    breaks = seq(0,max(Map_data$per_plot),10),
    labels = c(seq(0,max(Map_data$per_plot-10),10),">50")
    # option="magma",
    # direction = -1
  ) +
  scale_fill_viridis(
    name= "Changes in catch proportion \nand changing stocks (%)",
    limits = c(0,max(Map_data$per_plot)),
    breaks = seq(0,max(Map_data$per_plot),10),
    labels = c(seq(0,max(Map_data$per_plot-10),10),">50")
    # alpha = 0.8
    # option="magma",
    # direction = -1
  ) #+

ggsave(
  plot = Prop_map,
  filename = "./Figures/test.png",
  width = 10,
  height = 5,
  units = "in"
)

## ------------------- ##
# Regional Plots
## ------------------- ##

Lat_am_land <- World_map %>% 
  filter(region_wb %in% c("Latin America & Caribbean") |
           sau_fishing_entity %in% c("USA","Mexico")
  )

Lat_am_eez <- World_eez_sf %>% 
  select(sf_eez_name = eez_name,everything()) %>% 
  left_join(Matching_names) %>% 
  filter(
    !str_detect(sf_eez_name,"Pacific"),
    !sf_eez_name %in% c("USA (East Coast)","Suriname","Guyana","French Guiana","Bermuda (UK)","Brazil","El Salvador")
  )

Lat_am_data <- Map_data %>% 
  filter(source %in% Lat_am_eez$sf_eez_name)

Lat_am_paths <- paths %>% 
  filter(source %in% Lat_am_eez$sf_eez_name,
         target %in% Lat_am_eez$sf_eez_name)

Lat_am_map <- ggplot() +
  geom_sf(data = Lat_am_eez, size = 0.1, fill = "white", col = "gray30") +
  geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") +
  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
          aes(fill = avrg_spp_per),
          size = 0.1,col = "white") +
  geom_point(data = subset(Lat_am_data, time_step == time & per_plot < 0 ),
             aes(start_lon,
                 start_lat,
                 col = per_plot),
             size = 1,
             col = "gray30",
             shape = 20,
             stroke = 1.5) +
  geom_path(data = subset(Lat_am_paths, time_step == time & per_plot > 0),
            aes(X, Y,
                group = interaction(line_id,subline_id),
                col = per_plot
            ),
            size = 1,
            arrow = arrow(length = unit(0.05, "npc"),
                          ends = "first"),
            show.legend = NA) +
  coord_sf(xlim = c(-100, -55),ylim = c(9, 30)) + 
  labs(x = "", y = "") +
  scale_x_continuous(breaks = seq(-100,60,10)) +
  scale_y_continuous(breaks = seq(10,30,10)) +
  my_ggtheme_m(leg_pos = "bottom") +
  theme(legend.position = "",
        panel.background = element_rect(fill = "white", color = "white"),
        panel.grid.major =  element_line(color = "white")
  ) +
  scale_colour_viridis(
    name="Average Gain of\nStock-Share Ratio (%)",
    limits = c(0,max(Lat_am_data$per_plot)),
    breaks = seq(0,max(Lat_am_data$per_plot),10),
    # option="magma",
    # direction = -1
  )+
  scale_fill_viridis(
    name="Average Gain of\nStock-Share Ratio (%)",
    limits = c(0,max(Lat_am_data$per_plot)),
    breaks = seq(0,max(Lat_am_data$per_plot),10),
    # option="magma",
    # direction = -1
  )

## ------------------- ##
# The Med (Removed by Colette)
## ------------------- ##

# # Land polygon
# Med_land <- World_map %>% 
#   filter(region_wb %in% c("Europe & Central Asia","Middle East & North Africa") |
#            sau_fishing_entity %in% c("Spain","France")
#   )
# 
# # EEZ polygon
# Med_eez <- World_eez_sf %>% 
#   select(sf_eez_name = eez_name,everything()) %>% 
#   left_join(Matching_names) %>% 
#   filter(sau_fishing_entity %in% Med_land$sau_fishing_entity,
#          sf_eez_name != "Spain (Northwest)")
# 
# # Med data
# Med_data <- Map_data %>% 
#   filter(source %in% Med_eez$sf_eez_name)
# 
# # Med paths
# Med_paths <- paths %>% 
#   filter(source %in% Med_eez$sf_eez_name,
#          target %in% Med_eez$sf_eez_name)
# 
# # Med plot
# Med_map <- ggplot() +
#   geom_sf(data = Med_eez, size = 0.5, fill = "white", col = "gray30") +
#   geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
#   geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
#           aes(fill = avrg_spp_per),
#           size = 0.1,col = "white") +
#   geom_point(data = subset(Med_data, time_step == time & per_plot < 0 ),
#              aes(start_lon,
#                  start_lat,
#                  col = per_plot),
#              size = 1,
#              col = "gray30",
#              shape = 20,
#              stroke = 1.5) +
#   geom_path(data = subset(Med_paths, time_step == time & per_plot > 0),
#             aes(X, Y,
#                 group = interaction(line_id,subline_id),
#                 col = per_plot
#             ),
#             size = 1,
#             arrow = arrow(length = unit(0.05, "npc"),
#                           ends = "first"),
#             show.legend = NA) +
#   coord_sf(xlim = c(-3, 35),ylim = c(30, 45)) +
#   labs(x = "", y = "") +
#   scale_x_continuous(breaks = seq(-5,35,10)) +
#   scale_y_continuous(breaks = seq(30,50,5)) +
#   my_ggtheme_m() +
#   theme(legend.position = "",
#         panel.background = element_rect(fill = "white", color = "white"),
#         panel.grid.major =  element_line(color = "white")
#   ) +
#   scale_colour_viridis(
#     name="Average Gain of\nStock-Share Ratio (%)",
#     limits = c(0,max(Med_data$per_plot)),
#     breaks = seq(0,max(Med_data$per_plot),5),
#     # option="magma",
#     # direction = -1
#   )+
#   scale_fill_viridis(
#     name="Average Gain of\nStock-Share Ratio (%)",
#     limits = c(0,max(Med_data$per_plot)),
#     breaks = seq(0,max(Med_data$per_plot),5),
#     # option="magma",
#     # direction = -1
#   )#;Med_map

## ------------------- ##
# Northern Europe
## ------------------- ##

Eu_land <- World_map %>% 
  filter(
    continent == "Europe",
    sau_fishing_entity != "Iceland"
  )

# 
# ggplot(N_Eu_land) +
#   geom_sf()

Eu_eez <- World_eez_sf %>% 
  select(sf_eez_name = eez_name,everything()) %>% 
  left_join(Matching_names) %>% 
  filter(
    !sf_eez_name %in% c("Iceland", "Norway","Italy","Greece","Croatia","Bulgaria","Romania"),
    !str_detect(sf_eez_name,"Spain")
  )

# Eu_eez %>% filter(str_detect(sf_eez_name,"Romania"))

Eu_data <- Map_data %>% 
  filter(source %in% Eu_eez$sf_eez_name)

Eu_paths <- paths %>% 
  filter(source %in% Eu_eez$sf_eez_name,
         target %in% Eu_eez$sf_eez_name)

Eu_map <- ggplot() +
  geom_sf(data = Eu_eez, size = 0.5, fill = "white", col = "gray30") +
  geom_sf(data = grey_land, size = 0.1, fill = "grey50", col = "gray30") + # created for figure 2
  geom_sf(data = subset(n_taxa_map, sovereignt != "Antarctica" & time_step == time),
          aes(fill = avrg_spp_per),
          size = 0.1,col = "white") +
  geom_point(data = subset(Eu_data, time_step == time & per_plot < 0 ),
             aes(start_lon,
                 start_lat,
                 col = per_plot),
             size = 1,
             col = "gray30",
             shape = 20,
             stroke = 1.5) +
  geom_path(data = subset(Eu_paths, time_step == time & per_plot > 0),
            aes(X, Y,
                group = interaction(line_id,subline_id),
                col = per_plot
            ),
            size = 1,
            arrow = arrow(length = unit(0.05, "npc"),
                          ends = "first"),
            show.legend = NA) +
  coord_sf(xlim = c(-15, 30), ylim = c(45, 65)) +
  labs(x = "", y = "") +
  scale_y_continuous(breaks = seq(45,65,10)) +
  # scale_x_continuous(breaks = seq(45,65,5)) +
  my_ggtheme_m(leg pos = "bottom") +
  theme(legend.position = "",
        panel.background = element_rect(fill = "white", color = "white"),
        panel.grid.major =  element_line(color = "white")
  ) +
  scale_colour_viridis(
    name="Average Gain of\nStock-Share Ratio (%)",
    limits = c(0,max(Eu_data$per_plot)),
    breaks = seq(0,max(Eu_data$per_plot),5),
    # option="magma",
    # direction = -1
  )+
  scale_fill_viridis(
    name="Average Gain of\nStock-Share Ratio (%)",
    limits = c(0,max(Eu_data$per_plot)),
    breaks = seq(0,max(Eu_data$per_plot),5),
    # option="magma",
    # direction = -1
  )#;Eu_map

# Legend
map_legend <- get_legend(Prop_map)

Prop_map <- Prop_map + 
  theme(legend.position = "")

## ------------------- ##
# Render final plot
## ------------------- ##

# Clean memory
gc()

# Draw figure
plot_all <- ggdraw() +
  # Main map
  draw_plot(Prop_map, x = 0, y = 0.40, width = 1, height = 0.6) +
  # Latin America and the Car.
  draw_plot(Lat_am_map, x = 0.05, y = 0.00, width = 0.5, height = 0.50) +
  # Europe
  draw_plot(Eu_map, x = 0.53, y = 0.05, width = 0.35, height = 0.35) +
  # Legend
  draw_plot(map_legend, x = 0.25, y = 0, width = 0.59, height = 0.08) +
  # Labels
  draw_plot_label(label = c("A", "B", "The Caribbean","Northwestern Europe"),
                  size = c(20,20,15,15),
                  x = c(0.01,0.01,0.14,0.50),
                  y = c(1,0.45,0.44,0.44)
  )

name <- paste0("./Figures/props_",time,".png")

# Save plot
save_plot(
  name,
  plot_all,
  base_height = 10,
  base_width = 12
)

}
```